### 2025-02-09 products/manuscript/Manuscript.qmd
UpdateCode

---
title: "Example Manuscript Template for a Data Analysis Project"
author: "Asmith Joseph"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../../assets/dataanalysis-references.bib
csl: ../../assets/american-journal-of-epidemiology.csl
---

The structure below is one possible setup for a manuscript, or a general data analysis project (including the course project). Adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

#1.Load Required Libraries
install.packages("janitor")      # For cleaning column names
install.packages("lubridate")    # For working with dates

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
```

```{r}
# Load necessary libraries
library(tidyverse)
library(janitor)
library(lubridate)
library(readr)
library(here)

# Print the root directory to confirm the correct path
print(here())

# Read the CSV file using here()
PROMO_Data <- read_csv(here("PROMO_Data.csv"))

# View the dataset
View(PROMO_Data)

# View the data
View(PROMO_Data)

#3. Initial Data Inspection
# View structure of the dataset
str(PROMO_Data)

# View first few rows
head(PROMO_Data)

# Check column names
colnames(PROMO_Data)

# Get summary statistics
summary(PROMO_Data)

# Count missing values in each column
colSums(is.na(PROMO_Data))

```

**Authors**

* Asmith Joseph$^{1,2,*}$ (ORCID: 0009-0004-6875-0868)

**Author affiliations**

1. College of Public Health, University of Georgia, Athens, GA, USA.

$\land$ Corresponding author: asmith.joseph@uga.edu

$\dagger$ Disclaimer: The opinions expressed in this article are the
author's own and don't reflect their employer.

{{< pagebreak >}}

# Summary/Abstract
_Write a summary of your project._

{{< pagebreak >}}

# Introduction

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_

## Description of data and data source
_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._
The data for this analysis comes from the PROMOTE Birth Cohort 3 Randomized Controlled Trial, published in The Lancet (Kajubi et al., 2019). This was a double-blind, randomized, controlled, superiority trial conducted to compare monthly IPTp-SP vs. IPTp-DP among pregnant women in malaria-endemic regions.

## Questions/Hypotheses to be addressed
_State the research questions you plan to answer with this analysis._
 Does Malaria Preventive Treatment (DP vs. SP) Influence the Likelihood of Recurrent Placental Malaria in Subsequent Pregnancies?
Hypothesis: Women who received DP during pregnancy may have a lower likelihood of recurrent placental malaria in subsequent pregnancies compared to those who received SP, due to better efficacy of DP in preventing malaria infection.

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].

{{< pagebreak >}}

# Methods

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Schematic of workflow

Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it). @fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
We store those figures in the `assets` folder.

```{r}
#| label: fig-schematic
#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
#| echo: FALSE
knitr::include_graphics(here("assets","antigen-recognition.png"))
```

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._
_The dataset used in this analysis originates from the PROMOTE Birth Cohort 3 Randomized Controlled Trial, a double-blind, randomized controlled trial investigating the effectiveness of monthly intermittent preventive treatment of malaria (IPTp) in HIV-uninfected pregnant women and their infants. The trial compared sulfadoxine-pyrimethamine (SP) vs. dihydroartemisinin-piperaquine (DP) in malaria prevention.

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._
-A subset of variables was created from the ClinEpiDB Data Repository, including data from 782 participants. The dataset consists of 138 variables covering various aspects such as:
Study Arm, Participant Demographics, Clinical and Medical History, Obstetric History, Diagnosis and Obstetric Data (Pregnancy, Labor & Delivery, Delivery Complications), Medications Administered During Labor and Delivery, Aggregated Illness Data, Aggregated Lab Tests, Child Demographics
The dataset was downloaded as a text document and processed in RStudio to ensure it was clean and ready for analysis.

The cleaning process involved: Handling Missing Values, Renaming Columns for Consistency, Converting Data Types (e.g., categorical, numerical, date formats), and Removing Duplicate Records.

## Statistical analysis
_Explain anything related to your statistical analyses._
Planned Analysis 
Outcome & Predictors of Interest
Primary Outcome: Placental Malaria (Yes/No)
Key Predictor: Malaria Preventive Treatment (DP vs. SP)
Additional Factors to Consider:
Prior Malaria Exposure (Total malaria episodes before/during pregnancy)
Maternal Characteristics (Gravidity, Parity, Age)
Other Health Conditions (Hypertension, Anemia)

Statistical Plan
Descriptive Statistics: Compare PM rates across DP vs. SP groups.
Logistic Regression: Model the effect of DP vs. SP on the likelihood of PM, adjusting for potential confounders.
Interaction Analysis: Check if prior malaria exposure modifies the effect of DP vs. SP on PM risk.

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below. I generally recommend the `here` package.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/tables/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by gender."
#| echo: FALSE
knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like.

{{< pagebreak >}}

# References

### 2025-02-24 code/eda-code/MADAprojectcode.qmd
ProjectUpdate

---
title: "The Impact of Malaria Episodes and Treatment Regimens on Adverse Pregnancy Outcomes in Ugandan Women"
author: "Asmith Joseph"
date: "2025-02-23"
output: html_document
---

# Setup

```{r}
#load needed packages. make sure they are installed.
library(here) #for data loading/saving
library(dplyr)
library(skimr)
library(ggplot2)
library(tidyverse)
library(janitor)
library(knitr)
library(lubridate)
library(readr)
library(pwr)
library(glue)
```

```{r}
# Read the CSV file using here()
PROMO_Data <- read_csv(here("PROMO_Data.csv"))

# View the dataset
View(PROMO_Data)
```

```{r}
# Get the number of rows (observations) and columns (variables)
num_rows <- nrow(PROMO_Data)
num_columns <- ncol(PROMO_Data)

# Print the results
cat("Number of Observations (Rows):", num_rows, "\n")
cat("Number of Variables (Columns):", num_columns, "\n")

```

```{r}
# List all variable names
cat("List of Variables:\n")
print(colnames(PROMO_Data))
```

```{r}
# Display the data type of each variable
cat("Data Types of Each Variable:\n")
print(sapply(PROMO_Data, class))

```

```{r}

# Count of missing values in each variable
missing_values_per_column <- colSums(is.na(PROMO_Data))

# Display columns with missing data
cat("Missing Values per Variable:\n")
print(missing_values_per_column)
```

```{r}
#3. Initial Data Inspection
# View structure of the dataset
str(PROMO_Data)
```

```{r}
# View first few rows
head(PROMO_Data)
```

```{r}
# Check column names
colnames(PROMO_Data)
```

```{r}

# Get summary statistics
summary(PROMO_Data)
```

```{r}
# Count missing values in each column
colSums(is.na(PROMO_Data))

```

```{r}
# 3. Power analysis for comparing two independent groups (SP vs DP)
# Assuming a medium effect size (Cohen's d = 0.5)
# Alpha = 0.05, Power = 0.8

# Check available sample size using the correct dataframe
available_sample_size <- nrow(PROMO_Data)
cat("Available Sample Size:", available_sample_size, "\n")

# Check if the available sample size is sufficient
if (available_sample_size >= power_result$n) {
  cat("Sufficient sample size for detecting medium effect size.\n")
} else {
  cat("Sample size may be insufficient for detecting medium effect size.\n")
}
```

```{r}
# Define exposure and outcome variables
key_variables <- c(
  # Maternal health factors
  "Preeclampsia..HP_0100602.",
  "Postpartum.hemmorhage.after.spontaneous.vaginal.delivery..EUPATH_0042128.",
  "Hemorrhage.after.caesarean.section..EUPATH_0042084.",
  "Preterm.birth..ONTONEO_00000271.",
  "Infant.low.birth.weight...2500.g...EUPATH_0042398.",
  "Small.for.gestational.age..HP_0001518.",
  "Blood.transfusion..EUPATH_0042173.",

  # Infant health outcomes
  "Study.arm..EUPATH_0015457.",
  "Neonatal.death..HP_0003811."
)

```

```{r}
# Descriptive statistics for key continuous variables
library(dplyr)

# List of main numeric variables
numeric_vars <- c(
  "Total.malaria.episodes.during.pregnancy..EUPATH_0042235.",
  "Malaria.infection.rate.during.pregnancy..per.person.year...EUPATH_0042237."
)

# Generate summary statistics
numeric_summary <- PROMO_Data %>%
  select(all_of(numeric_vars)) %>%
  summary()

# Print summary
print(numeric_summary)

```

```{r}
# Descriptive stats for categorical variables
categorical_vars <- c(
  "Study.arm..EUPATH_0015457.",  # SP vs DP
  "Preterm.birth..ONTONEO_00000271.",
  "Stillbirth..HP_0003826.",
  "Infant.low.birth.weight...2500.g...EUPATH_0042398.",
  "Neonatal.death..HP_0003811.",
  "Placental.malaria..EUPATH_0042117."
)

# Count occurrences
categorical_summary <- PROMO_Data %>%
  select(all_of(categorical_vars)) %>%
  summarise(across(everything(), ~ sum(!is.na(.))))  # Non-missing counts

# Print counts
print(categorical_summary)

```

```{r}
# Histogram for malaria episodes during pregnancy
library(ggplot2)

ggplot(PROMO_Data, aes(x = Total.malaria.episodes.during.pregnancy..EUPATH_0042235.)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Distribution of Malaria Episodes During Pregnancy", x = "Number of Episodes", y = "Frequency")

```

```{r}
# Bar plot for treatment arms
ggplot(PROMO_Data, aes(x = Study.arm..EUPATH_0015457.)) +
  geom_bar(fill = "green", color = "black") +
  labs(title = "Distribution of Treatment Arms (SP vs. DP)", x = "Treatment Arm", y = "Count")

```

```{r}
# Boxplot for malaria infection rates across treatment arms
ggplot(PROMO_Data, aes(x = Study.arm..EUPATH_0015457., y = Malaria.infection.rate.during.pregnancy..per.person.year...EUPATH_0042237.)) +
  geom_boxplot(fill = "orange") +
  labs(title = "Malaria Infection Rate by Treatment Arm", x = "Treatment Arm", y = "Infection Rate (per person-year)")

```

```{r}
# Cross-tabulation of treatment arm vs. outcomes
cross_tab <- table(
  PROMO_Data$Study.arm..EUPATH_0015457.,
  PROMO_Data$Preterm.birth..ONTONEO_00000271.
)

# Print cross-tab
print(cross_tab)

```

```{r}
# Placental malaria vs. low birth weight
placental_vs_birth_weight <- table(
  PROMO_Data$Placental.malaria..EUPATH_0042117.,
  PROMO_Data$Infant.low.birth.weight...2500.g...EUPATH_0042398.
)

# Print the cross-tab
print(placental_vs_birth_weight)

```

```{r}
# Visualizing missing data
library(Amelia)

missmap(PROMO_Data, main = "Missing Data Map", col = c("red", "grey"), legend = TRUE)

```

```{r}
# Replace missing values with NA
PROMO_Data_clean <- PROMO_Data %>%
  filter(
    !is.na(Study.arm..EUPATH_0015457.),
    !is.na(Preterm.birth..ONTONEO_00000271.),
    !is.na(Stillbirth..HP_0003826.),
    !is.na(Infant.low.birth.weight...2500.g...EUPATH_0042398.)
  )

# Create a descriptive summary table
summary_table <- PROMO_Data_clean %>%
  select(
    Study.arm..EUPATH_0015457.,  # Treatment Arm (SP vs DP)
    Preterm.birth..ONTONEO_00000271.,  # Preterm Birth
    Stillbirth..HP_0003826.,  # Stillbirth
    Infant.low.birth.weight...2500.g...EUPATH_0042398.,  # Low Birth Weight
    Total.malaria.episodes.during.pregnancy..EUPATH_0042235.,  # Malaria Episodes
    Malaria.infection.rate.during.pregnancy..per.person.year...EUPATH_0042237.  # Infection Rate
  ) %>%
  tbl_summary(
    by = Study.arm..EUPATH_0015457.,  # Group by treatment arm
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",  # Mean and SD for continuous
      all_categorical() ~ "{n} / {N} ({p}%)"  # Count and percentage for categorical
    ),
    missing = "ifany"
  ) %>%
  add_overall() %>%
  modify_header(label = "**Characteristic**") %>%
  modify_spanning_header(all_stat_cols() ~ "**n/N (%) or mean (SD)**") %>%
  as_gt()  # Convert to GT table

```

```{r}
# Create a characteristics table
characteristics_table <- PROMO_Data %>%
  select(
    Study.arm..EUPATH_0015457.,  # Treatment Arm (SP vs DP)
    Preterm.birth..ONTONEO_00000271.,  # Preterm Birth
    Stillbirth..HP_0003826.,  # Stillbirth
    Infant.low.birth.weight...2500.g...EUPATH_0042398.,  # Low Birth Weight
    Total.malaria.episodes.during.pregnancy..EUPATH_0042235.,  # Malaria Episodes
    Malaria.infection.rate.during.pregnancy..per.person.year...EUPATH_0042237.  # Malaria Infection Rate
  ) %>%
  tbl_summary(
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",  # Mean (SD) for continuous variables
      all_categorical() ~ "{n} / {N} ({p}%)"  # Count and percentage for categorical variables
    ),
    missing = "ifany",  # Report missing values
    label = list(
      Study.arm..EUPATH_0015457. ~ "Treatment Arm (SP vs. DP)",
      Preterm.birth..ONTONEO_00000271. ~ "Preterm Birth",
      Stillbirth..HP_0003826. ~ "Stillbirth",
      Infant.low.birth.weight...2500.g...EUPATH_0042398. ~ "Low Birth Weight",
      Total.malaria.episodes.during.pregnancy..EUPATH_0042235. ~ "Number of Malaria Episodes",
      Malaria.infection.rate.during.pregnancy..per.person.year...EUPATH_0042237. ~ "Malaria Infection Rate"
    )
  ) %>%
  add_overall() %>%  # Add overall summary column
  modify_header(label = "**Characteristic**") %>%  # Customize column header
  modify_spanning_header(all_stat_cols() ~ "**n/N (%) or mean (SD)**") %>%  # Custom statistics label
  as_gt()  # Convert to a nicely formatted table using gt

```

Load the data.

```{r}
#Path to data. Note the use of the here() package and not absolute paths
data_location <- here::here("data","processed-data","processeddata.rds")
#load data
mydata <- readRDS(data_location)
```

# Data exploration through tables

Showing a bit of code to produce and save a summary table.

```{r}
summary_df = skimr::skim(mydata)
print(summary_df)
# save to file
summarytable_file = here("results", "tables", "summarytable.rds")
saveRDS(summary_df, file = summarytable_file)
```

We are saving the results to the `results` folder. Depending on how many tables/figures you have, it might make sense to have separate folders for each. And/or you could have separate folders for exploratory tables/figures and for final tables/figures. Just choose a setup that makes sense for your project and works for you, and provide enough documentation that someone can understand what you are doing.

# Data exploration through figures

Histogram plots for the continuous outcomes.

Height first.

```{r}
p1 <- mydata %>% ggplot(aes(x=Height)) + geom_histogram() 
plot(p1)
figure_file = here("results", "figures","height-distribution.png")
ggsave(filename = figure_file, plot=p1) 
```

Now weights.

```{r}
p2 <- mydata %>% ggplot(aes(x=Weight)) + geom_histogram() 
plot(p2)
figure_file = here("results", "figures","weight-distribution.png")
ggsave(filename = figure_file, plot=p2) 
```

Now height as function of weight.

```{r}
p3 <- mydata %>% ggplot(aes(x=Height, y=Weight)) + geom_point() + geom_smooth(method='lm')
plot(p3)
figure_file = here("results", "figures", "height-weight.png")
ggsave(filename = figure_file, plot=p3) 
```

Once more height as function of weight, stratified by gender. Note that there is so little data, it's a bit silly. But we'll plot it anyway.

```{r}
p4 <- mydata %>% ggplot(aes(x=Height, y=Weight, color = Gender)) + geom_point() + geom_smooth(method='lm')
plot(p4)
figure_file = here("results", "figures", "height-weight-stratified.png")
ggsave(filename = figure_file, plot=p4) 
```

# Notes

For your own explorations, tables and figures can be "quick and dirty". As long as you can see what's going on, there is no need to polish them. That's in contrast to figures you'll produce for your final products (paper, report, presentation, website, etc.). Those should look as nice, polished and easy to understand as possible.

### 2025-02-24 products/manuscript/Manuscript.qmd
P2UP
ProjectUpdate

@@
 ---
-title: "Example Manuscript Template for a Data Analysis Project"
+title: "The Impact of Malaria Episodes and Treatment Regimens on Adverse Pregnancy Outcomes in Ugandan Women"
 author: "Asmith Joseph"
@@
 The structure below is one possible setup for a manuscript, or a general data analysis project (including the course project). Adjust as needed.
+
 You don't need to have exactly these sections, but the content covering those sections should be addressed.
@@
 
-#1.Load Required Libraries
-install.packages("janitor")      # For cleaning column names
-install.packages("lubridate")    # For working with dates
-
-```{r, echo=FALSE, message=FALSE}
-# load a few R packages
-library(here)
-library(knitr)
-```
-
 ```{r}
-# Load necessary libraries
+# Loading necessary libraries
 library(tidyverse)
 library(janitor)
+library(knitr)
 library(lubridate)
@@
 library(here)
+```
 
+**Authors**
+
+-   Asmith Joseph$^{1*}$ (ORCID: 0009-0004-6875-0868)
+
+**Author affiliations**
+
+1.  College of Public Health, University of Georgia, Athens, GA, USA.
+
+$\land$ Corresponding author: asmith.joseph\@uga.edu
+
+$\dagger$ Disclaimer: The opinions expressed in this article are the author's own and don't reflect their employer.
+
+{{< pagebreak >}}
+
+# 1. Summary/Abstract
+
+*Write a summary of your project.*
+
+{{< pagebreak >}}
+
+# 2. Introduction
+
+## 2.1 General Background Information
+
+Malaria is a life-threatening disease caused by Plasmodium parasites, which are transmitted to humans through the bite of an infected female Anopheles mosquito (Talapko et al., 2019).
+According to the World Health Organization, in 2022, approximately 249 million malaria cases were reported across 85 endemic countries—an increase of 5 million cases compared to 2021.
+Notably, Uganda alone contributed more than 597,000 cases during that period (WHO, 2023).
+Despite significant efforts in prevention and treatment, malaria remains a major global health challenge.
+Vulnerable groups such as pregnant women, children, and immunocompromised individuals are at the highest risk, and Africa continues to bear the heaviest burden of this disease.
+Pregnant women with malaria tend to experience more severe clinical symptoms and poorer outcomes, including heightened risks of miscarriage, fetal loss, premature birth, low birth weight in newborns, and neonatal death (Caroline Lin Lin Chua et al., 2021).
+A multifaceted approach incorporating insecticide-treated bed nets (ITNs) and chemoprevention strategies has proven effective in mitigating the adverse health outcomes associated with malaria during pregnancy.
+Despite this progress, a 2017 report reveals that only 22% of pregnant women in Sub-Saharan Africa received an entire course of IPTp (Bauserman et al., 2019), underscoring a significant disparity in treatment access.
+Addressing this issue, the study seeks to: (a) examine the association between the frequency of malaria episodes during pregnancy and adverse birth outcomes, such as preterm birth and stillbirth; (b) assess whether infants born to mothers treated with DP are less likely to experience low birth weight compared to those treated with SP; and (c) explore the relationship between placental malaria and adverse birth outcomes, including preterm birth and low birth weight.
+
+## 2.1.1 Description of data and data source
+
+The dataset originates from a double-blind, randomized controlled trial evaluating monthly intermittent preventive treatment of malaria (IPTp) using either sulfadoxine-pyrimethamine (SP) or dihydroartemisinin-piperaquine (DP) in HIV-uninfected pregnant women and their infants.
+Data were rigorously collected throughout the trial, with detailed records capturing maternal health, pregnancy outcomes, and infant details.
+In cases of multiple births, separate records were created for each infant during postpartum visits, ensuring comprehensive outcome documentation.
+The dataset comprises 782 observations and an extensive array of variables, enabling advanced modeling of the impact of malaria episodes and treatment regimens on adverse pregnancy outcomes.
+
+The dataset encompasses several key variable categories.
+Maternal health and pregnancy outcomes are documented through unique participant identifiers, study arm assignment, and clinical indicators such as preeclampsia, various forms of postpartum hemorrhage (vaginal and cesarean), blood transfusions, laceration/episiotomy, congenital malformations, spontaneous abortions, stillbirths, and neonatal deaths, along with gestational parameters like the first day of the last menstrual period and estimated delivery dates.
+Reproductive history is captured via gravidity, parity, and counts of full-term births, preterm births, and abortions (both elective and spontaneous).
+Infant and child outcomes include details on delivery year, mode and location of delivery, birth complications, survival status, sex, and birth weight for up to seven children per participant.
+Malaria-specific measures record the total number of malaria episodes during pregnancy, infection rates per person-year, and results from routine blood smears, including the proportion of Plasmodium-positive.
+Additionally, the dataset contains information on maternal risk factors and comorbidities—such as education level, substance use, and various chronic conditions—and details on delivery complications and medications administered during labor.
+The data are publicly available through ClinEpiDB (Release #24, dated August 30, 2022).
+
+## 2.1.2 Questions/Hypotheses to be addressed
+
+The objective of this study is to examine the relationship between malaria exposure, treatment regimens, and adverse pregnancy outcomes in HIV-uninfected pregnant women receiving intermittent preventive treatment for malaria.
+Specifically, we will investigate whether an increased number of malaria episodes is associated with higher rates of preterm birth and stillbirth and whether the dihydroartemisinin-piperaquine (DP) regimen reduces these episodes compared to sulfadoxine-pyrimethamine (SP).
+We hypothesize that more malaria episodes correlate with an increased risk of adverse outcomes (Hypothesis 1) and that DP treatment results in fewer malaria episodes than SP (Hypothesis 2).
+
+In addition, the study will explore infant health outcomes and the role of placental malaria.
+We will test whether infants born to DP-treated mothers are less likely to have low birth weight and whether neonatal mortality rates differ between the treatment groups.
+We hypothesize that DP treatment will improve infant outcomes by reducing the incidence of low birth weight (Hypothesis 3) and that neonatal mortality rates will vary between the SP and DP arms (Hypothesis 4).
+Finally, we will assess whether placental malaria is associated with adverse birth outcomes, such as preterm birth and low birth weight, and if its prevalence differs by treatment regimen, with the expectation that placental malaria increases the risk of these outcomes (Hypothesis 5).
+
+{{< pagebreak >}}
+
+# 3. Methods
+
+#
+
+*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*
+
+## Schematic of workflow
+
+```{r}
+#| label: fig-schematic
+#| fig-cap: "Workflow schematic for analyzing associations between malaria during pregnancy, treatment regimens, and health outcomes."
+#| echo: FALSE
+
+library(DiagrammeR)
+
+grViz("
+  digraph statistical_workflow {
+    node [shape = rectangle, style = filled, fillcolor = lightblue]
+
+    Data_Collection [label = 'Data Collection']
+    Data_Cleaning [label = 'Data Cleaning & Preprocessing']
+    Exploratory_Analysis [label = 'Exploratory & Descriptive Analysis']
+    Statistical_Tests [label = 'Inferential Analysis\\n(Logistic, Poisson, Chi-Square)']
+    Sensitivity_Analyses [label = 'Sensitivity Analyses']
+    Results_Reporting [label = 'Reporting & Visualization']
+
+    Data_Collection -> Data_Cleaning -> Exploratory_Analysis -> Statistical_Tests -> Sensitivity_Analyses -> Results_Reporting
+  }
+")
+
+```
+
+## 3.1 Data aquisition
+
+The data for this study were obtained from ClinEpiDB, a public repository, specifically from Release #24 dated August 30, 2022.
+This dataset originates from a double-blind randomized controlled trial evaluating intermittent preventive treatment of malaria (IPTp) in HIV-uninfected pregnant women.
+The dataset, which includes 782 observations along with comprehensive maternal and infant health variables, was directly downloaded from the ClinEpiDB website without the need for authentication.
+Detailed metadata and documentation were also accessed via the platform, ensuring a clear understanding of variable definitions and study protocols.
+
+## 3.2 Data import and cleaning
+
+The dataset was imported into R directly from its source using rigorously tested and standardized data-loading functions.
+This approach ensured an accurate and complete transfer of information, providing you with confidence in the integrity of the data.
+We verified the dataset's structure and consistency upon import, confirming the presence of 782 observations and cross-validating variable names with the corresponding metadata.
+To ensure proper formatting for subsequent analyses, date fields were converted into Date objects, and categorical variables were transformed into factors, preserving data integrity and preventing modeling errors.
+
+Following the import and formatting, an initial inspection was conducted to assess the dataset's structure and content.
+This review involved examining internal structures using functions such as str(), summary(), colnames(), and head(), as well as generating descriptive statistics to understand each variable's central tendency, dispersion, and distribution.
+Visual inspections confirmed the data were correctly formatted and free from obvious entry errors, while exploratory visualizations identified potential outliers and anomalies.
+Recognizing the critical impact of missing data, we systematically quantified missing values by creating summary tables and using the missmap() function for visual assessment.
+Variables with more than 20% missing data were excluded to reduce potential bias.
+In contrast, those with less than 20% missing data were retained and appropriate imputation strategies were considered when necessary, ensuring that the final dataset was robust and reliable for subsequent statistical modeling.
+
+## 3.3 Statistical analysis
+
+# 3.3.0.1 Exploratory/Descriptive analyses
+
+We conducted comprehensive exploratory and descriptive analyses to gain an initial understanding of the dataset's structure, variable distributions, and data quality issues.
+In this phase, we summarized the demographic and clinical characteristics of the study population—including maternal age, parity, treatment allocation (SP vs.
+DP), and the distribution of malaria episodes during pregnancy—using descriptive statistics.
+Specifically, means, medians, standard deviations, and interquartile ranges were calculated for continuous variables, while frequencies and proportions were reported for categorical variables.
+Visualization techniques such as histograms, box plots, and bar charts helped identify outliers, skewed distributions, and other anomalies.
+
+In addition, cross-tabulations were performed to explore preliminary associations between key exposure and outcome variables, such as the frequency of low birth weight across treatment groups.
+We also assessed missing data patterns, using this information to guide decisions on data imputation or excluding variables with excessive missingness.
+Together, these initial analyses provided critical insights that guided model selection and ensured the validity of subsequent inferential statistical tests.
+
+*Logistic and Count Regression Analyses (Maternal Health Outcomes)* This section examines the relationship between malaria episodes during pregnancy and adverse outcomes like preterm birth and stillbirth.
+The exposure variable will be the total number of malaria episodes during pregnancy, with preterm birth and stillbirth as binary outcomes.
+Logistic regression will assess these associations, adjusting for confounders such as maternal age, parity, and socioeconomic status.
+Another analysis will investigate whether treatment type (SP vs.
+DP) reduces malaria episodes.
+Here, the exposure variable will be treatment, and the outcome will be the total number of malaria episodes.
+Poisson or negative binomial regression will be used to compare episode counts across treatment groups, adjusting for factors like baseline malaria status, geographic region, and seasonal variations.
+
+*Chi-Square Tests and Logistic Regression (Infant Health Outcomes)* This section focuses on birth weight, gestational age, and neonatal mortality outcomes.
+First, we will assess whether infants born to mothers treated with DP are less likely to have low birth weight compared to those treated with SP.
+A chi-square test will initially compare proportions, followed by logistic regression to adjust for confounders such as maternal nutrition and gestational age.
+Next, neonatal mortality will be compared between treatment groups using chi-square tests for unadjusted differences, with logistic regression models controlling for factors such as birth complications, preterm birth, and birth weight.
+Lastly, we will examine whether malaria infection rates during pregnancy are associated with low birth weight and preterm birth.
+Logistic regression models will be used for this analysis, adjusting for maternal age, parity, and gestational age.
+
+*Logistic Regression and Chi-Square Tests (Malaria and Pregnancy Outcomes)* This section will assess whether placental malaria is associated with preterm birth and low birth weight.
+Logistic regression models will evaluate these associations while adjusting for relevant confounders, such as the number of malaria episodes during pregnancy and maternal health variables.
+Additionally, the difference in placental malaria prevalence between SP and DP treatment groups will be examined using chi-square tests for initial comparisons and logistic regression for adjustments, controlling for factors like the number of malaria episodes and regional differences.
+
+{{< pagebreak >}}
+
+# 4. Results
+
+```{r}
 # Print the root directory to confirm the correct path
@@
 
-# View the data
-View(PROMO_Data)
+```
 
+```{r}
 #3. Initial Data Inspection
@@
 colSums(is.na(PROMO_Data))
-
 ```
 
-**Authors**
+## 4.1 Exploratory/Descriptive analysis
 
-* Asmith Joseph$^{1,2,*}$ (ORCID: 0009-0004-6875-0868)
-
-**Author affiliations**
-
-1. College of Public Health, University of Georgia, Athens, GA, USA.
-
-$\land$ Corresponding author: asmith.joseph@uga.edu
-
-$\dagger$ Disclaimer: The opinions expressed in this article are the
-author's own and don't reflect their employer.
-
-{{< pagebreak >}}
-
-# Summary/Abstract
-_Write a summary of your project._
-
-{{< pagebreak >}}
-
-# Introduction
-
-## General Background Information
-_Provide enough background on your topic that others can understand the why and how of your analysis_
-
-## Description of data and data source
-_Describe what the data is, what it contains, where it is from, etc. Eventually this might be part of a methods section._
-The data for this analysis comes from the PROMOTE Birth Cohort 3 Randomized Controlled Trial, published in The Lancet (Kajubi et al., 2019).
-This was a double-blind, randomized, controlled, superiority trial conducted to compare monthly IPTp-SP vs.
-IPTp-DP among pregnant women in malaria-endemic regions.
-
-## Questions/Hypotheses to be addressed
-_State the research questions you plan to answer with this analysis._
- Does Malaria Preventive Treatment (DP vs. SP) Influence the Likelihood of Recurrent Placental Malaria in Subsequent Pregnancies?
-Hypothesis: Women who received DP during pregnancy may have a lower likelihood of recurrent placental malaria in subsequent pregnancies compared to those who received SP, due to better efficacy of DP in preventing malaria infection.
-
-To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above and have the right bibtex key.
-Then you can include like this:
-
-Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a].
-
-{{< pagebreak >}}
-
-# Methods
-
-_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._
-
-## Schematic of workflow
-
-Sometimes you might want to show a schematic diagram/figure that was not created with code (if you can do it with code, do it).
-@fig-schematic is an example of some - completely random/unrelated - schematic that was generated with Biorender.
-We store those figures in the `assets` folder.
-
-```{r}
-#| label: fig-schematic
-#| fig-cap: "A figure that is manually generated and shows some overview/schematic. This has nothing to do with the data, it's just a random one from one of our projects I found and placed here."
-#| echo: FALSE
-knitr::include_graphics(here("assets","antigen-recognition.png"))
-```
-
-## Data aquisition
-_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._
-_The dataset used in this analysis originates from the PROMOTE Birth Cohort 3 Randomized Controlled Trial, a double-blind, randomized controlled trial investigating the effectiveness of monthly intermittent preventive treatment of malaria (IPTp) in HIV-uninfected pregnant women and their infants.
-The trial compared sulfadoxine-pyrimethamine (SP) vs.
-dihydroartemisinin-piperaquine (DP) in malaria prevention.
-
-## Data import and cleaning
-_Write code that reads in the file and cleans it so it's ready for analysis.
-Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts.
-If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g.
-as supplement in a paper).
-All materials, including files that contain code, should be commented well so everyone can follow along._
--A subset of variables was created from the ClinEpiDB Data Repository, including data from 782 participants. The dataset consists of 138 variables covering various aspects such as:
-Study Arm, Participant Demographics, Clinical and Medical History, Obstetric History, Diagnosis and Obstetric Data (Pregnancy, Labor & Delivery, Delivery Complications), Medications Administered During Labor and Delivery, Aggregated Illness Data, Aggregated Lab Tests, Child Demographics
-The dataset was downloaded as a text document and processed in RStudio to ensure it was clean and ready for analysis.
-
-The cleaning process involved: Handling Missing Values, Renaming Columns for Consistency, Converting Data Types (e.g., categorical, numerical, date formats), and Removing Duplicate Records.
-
-## Statistical analysis
-_Explain anything related to your statistical analyses._
-Planned Analysis 
-Outcome & Predictors of Interest
-Primary Outcome: Placental Malaria (Yes/No)
-Key Predictor: Malaria Preventive Treatment (DP vs. SP)
-Additional Factors to Consider:
-Prior Malaria Exposure (Total malaria episodes before/during pregnancy)
-Maternal Characteristics (Gravidity, Parity, Age)
-Other Health Conditions (Hypertension, Anemia)
-
-Statistical Plan
-Descriptive Statistics: Compare PM rates across DP vs. SP groups.
-Logistic Regression: Model the effect of DP vs. SP on the likelihood of PM, adjusting for potential confounders.
-Interaction Analysis: Check if prior malaria exposure modifies the effect of DP vs. SP on PM risk.
-
-{{< pagebreak >}}
-
-# Results
-
-## Exploratory/Descriptive analysis
-
-_Use a combination of text/tables/figures to explore and describe your data.
+*Use a combination of text/tables/figures to explore and describe your data.
 Show the most important descriptive results here.
 Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project._
+Even more can be in the R and Quarto files that are part of your project.*
 
@@
 
-## Basic statistical analysis
+## 4.2 Basic statistical analysis
 
-_To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
+*To get some further insight into your data, if reasonable you could compute simple statistics (e.g.
 simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable.
-Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._
+Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*
 
@@
 
-## Full analysis
+## 4.3 Full analysis
 
-_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
+*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc.
 This might again be code that is best placed in one or several separate R scripts that need to be well documented.
 You want the code to produce figures and data ready for display as tables, and save those.
-Then you load them here._
+Then you load them here.*
 
@@
 
-# Discussion
+# 5. Discussion
 
 ## Summary and Interpretation
-_Summarize what you did, what you found and what it means._
+
+*Summarize what you did, what you found and what it means.*
 
 ## Strengths and Limitations
-_Discuss what you perceive as strengths and limitations of your analysis._
+
+*Discuss what you perceive as strengths and limitations of your analysis.*
 
 ## Conclusions
-_What are the main take-home messages?_
 
-_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_
+*What are the main take-home messages?*
+
+*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*
 

### 2025-02-24 products/manuscript/Manuscript.qmd
Docx

@@
     highlight-style: github
+    prefer-html: true  # Only here
 bibliography: ../../assets/dataanalysis-references.bib

### 2025-02-24 code/MADAprojectcode.qmd
Doc

--- 2025-02-24 code/eda-code/MADAprojectcode.qmd
@@
 
+```{r}
+#| label: fig-schematic
+#| fig-cap: "Workflow schematic for analyzing associations between malaria during pregnancy, treatment regimens, and health outcomes."
+#| echo: FALSE
+#| screenshot.force: TRUE  # This forces a screenshot to be taken
+
+library(DiagrammeR)
+
+grViz("
+  digraph statistical_workflow {
+    node [shape = rectangle, style = filled, fillcolor = lightblue]
+
+    Data_Collection [label = 'Data Collection']
+    Data_Cleaning [label = 'Data Cleaning & Preprocessing']
+    Exploratory_Analysis [label = 'Exploratory & Descriptive Analysis']
+    Statistical_Tests [label = 'Inferential Analysis\\n(Logistic, Poisson, Chi-Square)']
+    Sensitivity_Analyses [label = 'Sensitivity Analyses']
+    Results_Reporting [label = 'Reporting & Visualization']
+
+    Data_Collection -> Data_Cleaning -> Exploratory_Analysis -> Statistical_Tests -> Sensitivity_Analyses -> Results_Reporting
+  }
+")
+
+```
+
+```
+
 Load the data.

### 2025-02-24 products/manuscript/Manuscript.qmd
Doc

--- 2025-02-24 products/manuscript/Manuscript.qmd
@@
     highlight-style: github
+    prefer-html: true
+    reference-location: margin  # Enable reference links
+knitr:
+  opts_chunk:
+    screenshot.force: TRUE
 bibliography: ../../assets/dataanalysis-references.bib
@@
 ---
-
-The structure below is one possible setup for a manuscript, or a general data analysis project (including the course project). Adjust as needed.
-
-You don't need to have exactly these sections, but the content covering those sections should be addressed.
-
-This uses MS Word as output format.
-[See here](https://quarto.org/docs/output-formats/ms-word.html) for more information.
-You can switch to other formats, like html or pdf.
-See [the Quarto documentation](https://quarto.org/) for other formats.
-
-```{r}
-# Loading necessary libraries
-library(tidyverse)
-library(janitor)
-library(knitr)
-library(lubridate)
-library(readr)
-library(here)
-```
 
@@
 
-```{r}
-#| label: fig-schematic
-#| fig-cap: "Workflow schematic for analyzing associations between malaria during pregnancy, treatment regimens, and health outcomes."
-#| echo: FALSE
-
-library(DiagrammeR)
-
-grViz("
-  digraph statistical_workflow {
-    node [shape = rectangle, style = filled, fillcolor = lightblue]
-
-    Data_Collection [label = 'Data Collection']
-    Data_Cleaning [label = 'Data Cleaning & Preprocessing']
-    Exploratory_Analysis [label = 'Exploratory & Descriptive Analysis']
-    Statistical_Tests [label = 'Inferential Analysis\\n(Logistic, Poisson, Chi-Square)']
-    Sensitivity_Analyses [label = 'Sensitivity Analyses']
-    Results_Reporting [label = 'Reporting & Visualization']
-
-    Data_Collection -> Data_Cleaning -> Exploratory_Analysis -> Statistical_Tests -> Sensitivity_Analyses -> Results_Reporting
-  }
-")
-
-```
-
 ## 3.1 Data aquisition
@@
 
-```{r}
-# Print the root directory to confirm the correct path
-print(here())
-
-# Read the CSV file using here()
-PROMO_Data <- read_csv(here("PROMO_Data.csv"))
-
-# View the dataset
-View(PROMO_Data)
-
-```
-
-```{r}
-#3. Initial Data Inspection
-# View structure of the dataset
-str(PROMO_Data)
-
-# View first few rows
-head(PROMO_Data)
-
-# Check column names
-colnames(PROMO_Data)
-
-# Get summary statistics
-summary(PROMO_Data)
-
-# Count missing values in each column
-colSums(is.na(PROMO_Data))
-```
-
 ## 4.1 Exploratory/Descriptive analysis
-
-*Use a combination of text/tables/figures to explore and describe your data.
-Show the most important descriptive results here.
-Additional ones should go in the supplement.
-Even more can be in the R and Quarto files that are part of your project.*
-
-@tbl-summarytable shows a summary of the data.
 
@@
 
-```{r}
-#| label: tbl-summarytable
-#| tbl-cap: "Data summary table."
-#| echo: FALSE
-resulttable=readRDS("../../results/tables/summarytable.rds")
-knitr::kable(resulttable)
-```
-
 ## 4.2 Basic statistical analysis
@@
 Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*
-
-@fig-result shows a scatterplot figure produced by one of the R scripts.
-
-```{r}
-#| label: fig-result
-#| fig-cap: "Height and weight stratified by gender."
-#| echo: FALSE
-knitr::include_graphics(here("results","figures","height-weight-stratified.png"))
-```
 
@@
 Then you load them here.*
-
-Example @tbl-resulttable2 shows a summary of a linear model fit.
-
-```{r}
-#| label: tbl-resulttable2
-#| tbl-cap: "Linear model fit table."
-#| echo: FALSE
-resulttable2 = readRDS(here("results","tables","resulttable2.rds"))
-knitr::kable(resulttable2)
-```
 

### 2025-02-26 products/manuscript/Manuscript.qmd
New Folder

@@
 
-# 1. Summary/Abstract
+# Summary/Abstract
 
@@
 
-# 2. Introduction
+# Introduction
 
-## 2.1 General Background Information
+## General Background Information
 
@@
 
-## 2.1.1 Description of data and data source
+## Description of data and data source
 
@@
 
-## 2.1.2 Questions/Hypotheses to be addressed
+## Questions/Hypotheses to be addressed
 
@@
 
-# 3. Methods
-
-#
+# Methods
 
@@
 
-## 3.1 Data aquisition
+## Data aquisition
 
@@
 
-## 3.2 Data import and cleaning
+## Data import and cleaning
 
@@
 
-## 3.3 Statistical analysis
+## Statistical analysis
 
-# 3.3.0.1 Exploratory/Descriptive analyses
+# Exploratory/Descriptive analyses
 
@@
 
-*Logistic and Count Regression Analyses (Maternal Health Outcomes)* This section examines the relationship between malaria episodes during pregnancy and adverse outcomes like preterm birth and stillbirth.
+Logistic and Count Regression Analyses (Maternal Health Outcomes)
+
+This section examines the relationship between malaria episodes during pregnancy and adverse outcomes like preterm birth and stillbirth.
 The exposure variable will be the total number of malaria episodes during pregnancy, with preterm birth and stillbirth as binary outcomes.
@@
 
-*Chi-Square Tests and Logistic Regression (Infant Health Outcomes)* This section focuses on birth weight, gestational age, and neonatal mortality outcomes.
+Chi-Square Tests and Logistic Regression (Infant Health Outcomes)
+
+This section focuses on birth weight, gestational age, and neonatal mortality outcomes.
 First, we will assess whether infants born to mothers treated with DP are less likely to have low birth weight compared to those treated with SP.
@@
 
-*Logistic Regression and Chi-Square Tests (Malaria and Pregnancy Outcomes)* This section will assess whether placental malaria is associated with preterm birth and low birth weight.
+Logistic Regression and Chi-Square Tests (Malaria and Pregnancy Outcomes)
+
+This section will assess whether placental malaria is associated with preterm birth and low birth weight.
 Logistic regression models will evaluate these associations while adjusting for relevant confounders, such as the number of malaria episodes during pregnancy and maternal health variables.
@@
 
-# 4. Results
+# Results
 
-## 4.1 Exploratory/Descriptive analysis
+## Exploratory/Descriptive analysis
 
@@
 
-## 4.2 Basic statistical analysis
+## Basic statistical analysis
 
@@
 
-## 4.3 Full analysis
+## Full analysis
 
@@
 
-# 5. Discussion
+# Discussion
 

### 2025-03-11 code/eda-code/eda_MADAprojectcode.qmd
Project Update

---
title: "The Impact of Malaria Episodes and Treatment Regimens on Adverse Pregnancy Outcomes in Ugandan Women"
author: "Asmith Joseph"
date: "2025-02-23"
output: html_document
---

```{r}
# **1. Load Required Libraries**
# Essential libraries for data handling, visualization, and analysis

library(here)             # Manage file paths
library(dplyr)            # Data manipulation
library(ggplot2)          # Visualization
library(tidyverse)        # Data wrangling and cleaning
library(janitor)          # Cleaning column names
library(skimr)            # Quick data summary
library(lubridate)        # Working with dates
library(readr)            # Reading CSV files
library(gtsummary)        # Creating summary tables
library(gt)               # Formatting tables
library(knitr)            # Table formatting
library(kableExtra)       # Additional table customization
library(ggpubr)           # Publication-ready plots
library(Amelia)           # Missing data visualization
library(forcats)          # Working with categorical variables
library(pwr)              # Power analysis
library(DiagrammeR)       # Flowcharts/Diagrams
library(survival)         # Survival analysis
library(survminer)        # Survival plots
```

# Data Import & Initial Inspection 
```{r}
# Read the dataset
PROMO_Data <- read_csv(here("data", "raw-data", "PROMO_Data.csv"))

# **Dataset Overview**
cat("Number of Observations:", nrow(PROMO_Data), "\n")
cat("Number of Variables:", ncol(PROMO_Data), "\n")

# **Inspect Data Structure**
str(PROMO_Data)
head(PROMO_Data)

# **Check Missing Values**
missing_values <- colSums(is.na(PROMO_Data))
print(missing_values)

# **Summary Statistics**
summary(PROMO_Data)

```

# Data Cleaning & Transformation
```{r}
# Clean Column Names**
PROMO_Data <- PROMO_Data %>% clean_names()

# **Remove Empty Rows**
PROMO_Data <- PROMO_Data %>%
  filter(!is.na(study_arm), !is.na(stillbirth), !is.na(neonatal_death), !is.na(infant_low_birth_weight_2500_g))

# **Handle Missing Values**
PROMO_Data <- PROMO_Data %>%
  mutate(across(where(is.numeric), ~ ifelse(is.na(.), median(., na.rm = TRUE), .))) %>%
  mutate(across(where(is.character), ~ replace_na(., "Unknown")))

# **Convert Data Types**
PROMO_Data <- PROMO_Data %>%
  mutate(across(c(study_arm, stillbirth, neonatal_death, placental_malaria, delivery_mode), as.factor)) %>%
  mutate(across(contains("date"), as.Date, format="%Y-%m-%d"))

# **Verify Changes**
str(PROMO_Data)

```

# Exploratory Data Analysis (EDA)

```{r}
# **Summary of Numeric Variables**
skim(PROMO_Data)

# **Frequency Counts of Key Categorical Variables**
PROMO_Data %>%
  count(study_arm, sort = TRUE)

PROMO_Data %>%
  count(placental_malaria, sort = TRUE)

PROMO_Data %>%
  count(preterm_birth_ontoneo, sort = TRUE)

```

# Characteristics Tables
*Baseline Characteristics Table*

```{r}
# Select key variables for baseline characteristics
baseline_vars <- c("study_arm", "education_level_omrse", "gravidity", "parity", 
                   "placental_malaria", "preterm_birth_ontoneo", "stillbirth", 
                   "neonatal_death", "infant_low_birth_weight_2500_g")

# **Ensure selected variables exist**
baseline_vars <- baseline_vars[baseline_vars %in% colnames(PROMO_Data)]

# Generate formatted table using gtsummary
baseline_characteristics <- PROMO_Data %>%
  select(all_of(baseline_vars)) %>%
  tbl_summary(by = study_arm) %>%
  add_p() %>%
  modify_header(label = "**Variable**") %>%
  bold_labels()

baseline_characteristics

```

*Continuous Characteristics by Study Arm*

```{r}
# Function to summarize continuous variables
summarize_continuous <- function(var) {
  sp_mean <- mean(PROMO_Data[[var]][PROMO_Data$study_arm == "Sulphadoxine-pyrimethamine (SP)"], na.rm = TRUE)
  sp_sd <- sd(PROMO_Data[[var]][PROMO_Data$study_arm == "Sulphadoxine-pyrimethamine (SP)"], na.rm = TRUE)

  dp_mean <- mean(PROMO_Data[[var]][PROMO_Data$study_arm == "Dihydroartemisinin-piperaquine (DP)"], na.rm = TRUE)
  dp_sd <- sd(PROMO_Data[[var]][PROMO_Data$study_arm == "Dihydroartemisinin-piperaquine (DP)"], na.rm = TRUE)

  p_value <- t.test(PROMO_Data[[var]] ~ PROMO_Data$study_arm, var.equal = TRUE, na.rm = TRUE)$p.value

  return(c(sprintf("%.1f (%.1f)", sp_mean, sp_sd),
           sprintf("%.1f (%.1f)", dp_mean, dp_sd),
           sprintf("%.3f", p_value)))
}

# Create summary table
summary_table <- tibble(
  Characteristic = c("Maternal Age (years)", "Gravidity", "Gestational Age at Delivery (weeks)", "Malaria Episodes During Pregnancy"),
  `SP (Mean ± SD)` = sapply(c("age_at_enrollment_years", "gravidity", "gestational_age_at_delivery_weeks", "total_malaria_episodes_during_pregnancy"), summarize_continuous)[1,],
  `DP (Mean ± SD)` = sapply(c("age_at_enrollment_years", "gravidity", "gestational_age_at_delivery_weeks", "total_malaria_episodes_during_pregnancy"), summarize_continuous)[2,],
  `p-value` = sapply(c("age_at_enrollment_years", "gravidity", "gestational_age_at_delivery_weeks", "total_malaria_episodes_during_pregnancy"), summarize_continuous)[3,]
)

# Display table using gt
summary_table %>%
  gt() %>%
  tab_header(title = md("**Table: Continuous Characteristics by Study Arm**")) %>%
  cols_align(align = "center") %>%
  tab_options(
    table.font.size = px(12),
    heading.title.font.size = px(14),
    table.border.top.style = "solid",
    table.border.bottom.style = "solid"
  )

```

```{r}
# Load required packages
library(dplyr)
library(gtsummary)

# Select relevant variables for malaria episodes and birth outcomes
malaria_outcomes_table <- PROMO_Data %>%
  select(study_arm, 
         total_malaria_episodes_during_pregnancy, 
         placental_malaria, 
         preterm_birth_ontoneo, 
         stillbirth, 
         infant_low_birth_weight_2500_g) %>%
  tbl_summary(
    by = study_arm,  # Compare DP vs. SP
    statistic = list(
      total_malaria_episodes_during_pregnancy ~ "{mean} ± {sd}",
      all_categorical() ~ "{n} ({p}%)"
    ),
    label = list(
      total_malaria_episodes_during_pregnancy ~ "Total Malaria Episodes (Mean ± SD)",
      placental_malaria ~ "Placental Malaria (Histology)",
      preterm_birth_ontoneo ~ "Preterm Birth",
      stillbirth ~ "Stillbirth",
      infant_low_birth_weight_2500_g ~ "Low Birth Weight (<2500g)"
    )
  ) %>%
  add_p() %>%
  bold_labels()

# Print the formatted table
malaria_outcomes_table

```

*Categorical Characteristics by Study Arm*

```{r}
# Generate a summary table for categorical characteristics
categorical_table <- PROMO_Data %>%
  select(study_arm, stillbirth, neonatal_death, infant_low_birth_weight_2500_g) %>%
  tbl_summary(by = study_arm, 
              statistic = list(all_categorical() ~ "{n} ({p}%)"),
              label = list(stillbirth ~ "Stillbirth",
                           neonatal_death ~ "Neonatal Death",
                           infant_low_birth_weight_2500_g ~ "Low Birth Weight")) %>%
  add_p() %>%
  bold_labels()

categorical_table

```

# Data Visualization

*Histogram: Malaria Episodes*

```{r}
ggplot(PROMO_Data, aes(x = total_malaria_episodes_during_pregnancy)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black", alpha = 0.7) +
  labs(title = "Distribution of Malaria Episodes", x = "Malaria Episodes", y = "Count") +
  theme_minimal()

```

*Boxplot: Gestational Age at Delivery*
```{r}
ggplot(PROMO_Data, aes(x = study_arm, y = gestational_age_at_delivery_weeks, fill = study_arm)) +
  geom_boxplot(alpha = 0.7) +
  labs(title = "Gestational Age at Delivery by Study Arm",
       x = "Study Arm", y = "Gestational Age (weeks)") +
  theme_minimal() +
  theme(legend.position = "none")

```

*Placental Malaria Prevalence by Study Arm*
```{r}
ggplot(PROMO_Data, aes(x = placental_malaria, fill = study_arm)) +
  geom_bar(position = "dodge", alpha = 0.7) +
  labs(title = "Placental Malaria Prevalence by Study Arm",
       x = "Placental Malaria", y = "Count") +
  theme_minimal()

```

*Stillbirth & Neonatal Death by Study Arm*
```{r}
ggplot(PROMO_Data, aes(x = study_arm, fill = stillbirth)) +
  geom_bar(position = "dodge", alpha = 0.7) +
  labs(title = "Stillbirth Rates by Study Arm", x = "Study Arm", y = "Count") +
  theme_minimal()
```

```{r}
ggplot(PROMO_Data, aes(x = study_arm, fill = neonatal_death)) +
  geom_bar(position = "dodge", alpha = 0.7) +
  labs(title = "Neonatal Mortality by Study Arm", x = "Study Arm", y = "Count") +
  theme_minimal()

```

# Statistical Analysis

*Chi-Square Test: Study Arm vs. Stillbirth*
```{r}
table_study_stillbirth <- table(PROMO_Data$study_arm, PROMO_Data$stillbirth)
chisq.test(table_study_stillbirth)

```

*Logistic Regression: Neonatal Death Prediction*
```{r}
logit_neonatal_death <- glm(neonatal_death ~ study_arm + total_malaria_episodes_during_pregnancy + gestational_age_at_delivery_weeks, 
                            data = PROMO_Data, family = binomial)

summary(logit_neonatal_death)

```

*Directed Acyclic Graph (DAG)*
```{r}
library(ggdag)
library(dagitty)

dag <- dagitty("dag {
  'Malaria Exposure' -> 'Placental Malaria'
  'Placental Malaria' -> 'Inflammation'
  'Inflammation' -> 'Pregnancy Outcomes'
  'Malaria Exposure' -> 'Pregnancy Outcomes'
  'Pregnancy Outcomes' -> 'Neonatal Health'
  'Treatment (SP vs. DP)' -> 'Malaria Exposure'
  'Treatment (SP vs. DP)' -> 'Pregnancy Outcomes'
}")

ggdag(dag, text = TRUE, use_labels = "name") +
  theme_minimal() +
  ggtitle("Directed Acyclic Graph (DAG) of Malaria in Pregnancy and Birth Outcomes")

```

*Kaplan-Meier Survival Curve*
```{r}
PROMO_Data$surv_obj <- Surv(PROMO_Data$gestational_age_at_enrollment_weeks, 
                             PROMO_Data$total_malaria_episodes_during_pregnancy > 0)

km_fit <- survfit(surv_obj ~ study_arm, data = PROMO_Data)

ggsurvplot(km_fit, data = PROMO_Data,
           conf.int = TRUE, pval = TRUE, risk.table = TRUE,
           legend.title = "Treatment Arm",
           ggtheme = theme_minimal(),
           title = "Time to Malaria Infection During Pregnancy")

```

### 2025-03-11 products/manuscript/Manuscript.qmd
Project Update

@@
 ---
+knitr:
+  opts_chunk:
+    echo: false
+    message: false
+    warning: false
+
 title: "The Impact of Malaria Episodes and Treatment Regimens on Adverse Pregnancy Outcomes in Ugandan Women"
@@
     number-sections: true
-    highlight-style: github
-    prefer-html: true
-    reference-location: margin  # Enable reference links
-knitr:
-  opts_chunk:
-    screenshot.force: TRUE
+    reference-location: margin
 bibliography: ../../assets/dataanalysis-references.bib
@@
 ---
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# Load Required Libraries
+# Essential libraries for data handling, visualization, and analysis
+
+library(here)             # Manage file paths
+library(tidyverse)        # Core data wrangling: includes dplyr, ggplot2, readr, forcats, lubridate
+library(janitor)          # Cleaning column names
+library(skimr)            # Quick data summary
+library(gtsummary)        # Creating summary tables
+library(gt)               # Formatting tables
+library(knitr)            # Table formatting
+library(kableExtra)       # Additional table customization
+library(ggpubr)           # Publication-ready plots
+library(Amelia)           # Missing data visualization
+library(pwr)              # Power analysis
+library(DiagrammeR)       # Flowcharts/Diagrams
+library(survival)         # Survival analysis
+library(survminer)        # Survival plots
+
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# Define the correct file path using the `here` package
+data_path <- here("data", "raw-data", "PROMO_Data.csv")
+
+# Read the CSV file, suppressing messages
+PROMO_Data <- read_csv(data_path, show_col_types = FALSE)
+```
 
@@
 
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# Data Import & Initial Inspection
+
+# Read the dataset
+PROMO_Data <- read_csv(here("data", "raw-data", "PROMO_Data.csv"), show_col_types = FALSE)
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+#  Data Cleaning & Transformation
+
+# Load required libraries (if not already loaded)
+library(janitor)   # For clean_names()
+
+# Clean Column Names
+PROMO_Data <- PROMO_Data %>% clean_names()
+
+# Remove Empty Rows
+PROMO_Data <- PROMO_Data %>%
+  filter(!is.na(study_arm), !is.na(stillbirth), !is.na(neonatal_death), !is.na(infant_low_birth_weight_2500_g))
+
+#  Handle Missing Values
+PROMO_Data <- PROMO_Data %>%
+  replace_na(list(
+    study_arm = "Unknown",
+    stillbirth = "Unknown",
+    neonatal_death = "Unknown",
+    placental_malaria = "Unknown",
+    delivery_mode = "Unknown"
+  )) %>%
+  mutate(across(where(is.numeric), ~ ifelse(is.na(.), median(., na.rm = TRUE), .)))
+
+#  Convert Data Types
+PROMO_Data <- PROMO_Data %>%
+  mutate(across(c(study_arm, stillbirth, neonatal_death, placental_malaria, delivery_mode), as.factor)) %>%
+  mutate(across(contains("date"), as.Date, format="%Y-%m-%d"))
+
+#  Verify Changes
+glimpse(PROMO_Data)  # More readable than str()
+
+#  Summary of Numeric Variables
+print(skim(PROMO_Data))  # Explicitly print skim output
+
+#  Frequency Counts of Key Categorical Variables
+PROMO_Data %>% count(study_arm, sort = TRUE) %>% print()
+PROMO_Data %>% count(placental_malaria, sort = TRUE) %>% print()
+PROMO_Data %>% count(preterm_birth_ontoneo, sort = TRUE) %>% print()
+
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+#  Load required libraries
+library(dplyr)
+library(gtsummary)
+
+#  Select key baseline variables (ensure they exist)
+baseline_vars <- c("study_arm", "education_level_omrse", "gravidity", "parity", 
+                   "placental_malaria", "preterm_birth_ontoneo", "stillbirth", 
+                   "neonatal_death", "infant_low_birth_weight_2500_g")
+
+#  Ensure only existing variables are used
+baseline_vars <- intersect(baseline_vars, colnames(PROMO_Data))
+
+# Generate table only if dataset has rows
+if (nrow(PROMO_Data) > 0) {
+  baseline_characteristics <- PROMO_Data %>%
+    select(all_of(baseline_vars)) %>%
+    tbl_summary(by = study_arm) %>%
+    add_p() %>%
+    modify_header(label = "**Variable**") %>%
+    bold_labels()
+
+  # Display only the formatted table
+  baseline_characteristics
+} else {
+  cat("️ No data available for analysis.")
+}
+
+```
+
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
+# Load required libraries
+library(dplyr)
+library(gtsummary)
+
+# Define key variables (check if they exist in dataset)
+malaria_outcome_vars <- c("study_arm", 
+                           "total_malaria_episodes_during_pregnancy", 
+                           "placental_malaria", 
+                           "preterm_birth_ontoneo", 
+                           "stillbirth", 
+                           "infant_low_birth_weight_2500_g")
+
+#  Keep only variables that exist in dataset
+malaria_outcome_vars <- intersect(malaria_outcome_vars, colnames(PROMO_Data))
+
+#  Generate the table only if dataset has rows
+if (nrow(PROMO_Data) > 0) {
+  malaria_outcomes_table <- PROMO_Data %>%
+    select(all_of(malaria_outcome_vars)) %>%
+    tbl_summary(
+      by = study_arm,  # Compare DP vs. SP groups
+      statistic = list(
+        total_malaria_episodes_during_pregnancy ~ "{mean} ± {sd}",  # Show mean ± SD
+        all_categorical() ~ "{n} ({p}%)"  # Show counts and percentages for categorical variables
+      ),
+      label = list(
+        total_malaria_episodes_during_pregnancy ~ "Total Malaria Episodes (Mean ± SD)",
+        placental_malaria ~ "Placental Malaria (Histology)",
+        preterm_birth_ontoneo ~ "Preterm Birth",
+        stillbirth ~ "Stillbirth",
+        infant_low_birth_weight_2500_g ~ "Low Birth Weight (<2500g)"
+      )
+    ) %>%
+    add_p() %>%  # Add p-values for comparisons
+    modify_header(label = "**Variable**") %>%  # Ensure readable column headers
+    bold_labels()  # Bold table labels for clarity
+
+  #  Display only the formatted table
+  malaria_outcomes_table
+} else {
+  cat("️ No data available for analysis.")
+}
+
+```
+
 ## Exploratory/Descriptive analysis

### 2025-03-12 code/eda-code/eda_MADAprojectcode.qmd
Project Update

@@
 ```{r}
-# Load required packages
+colnames(PROMO_Data)
+
+```
+
+```{r}
+# Load required libraries
+# Load required libraries
 library(dplyr)
@@
 
-# Select relevant variables for malaria episodes and birth outcomes
-malaria_outcomes_table <- PROMO_Data %>%
-  select(study_arm, 
-         total_malaria_episodes_during_pregnancy, 
-         placental_malaria, 
-         preterm_birth_ontoneo, 
-         stillbirth, 
-         infant_low_birth_weight_2500_g) %>%
-  tbl_summary(
-    by = study_arm,  # Compare DP vs. SP
-    statistic = list(
-      total_malaria_episodes_during_pregnancy ~ "{mean} ± {sd}",
-      all_categorical() ~ "{n} ({p}%)"
-    ),
-    label = list(
+# Define malaria outcome variables
+malaria_outcome_vars <- c(
+  "total_malaria_episodes_during_pregnancy",
+  "placental_malaria",
+  "preterm_birth_ontoneo",
+  "stillbirth",
+  "infant_low_birth_weight_2500_g"
+)
+
+# Ensure that `study_arm` is in the dataset
+if (!"study_arm" %in% colnames(PROMO_Data)) {
+  stop("⚠️ Error: `study_arm` column not found in PROMO_Data. Please check column names.")
+}
+
+# Remove `stillbirth` and any other variable that has only one level
+valid_outcomes <- malaria_outcome_vars[sapply(PROMO_Data[malaria_outcome_vars], function(x) length(unique(na.omit(x))) > 1)]
+
+# Ensure `study_arm` is included in the selected variables
+valid_outcomes <- c(valid_outcomes, "study_arm")
+
+# Define label mappings dynamically based on valid variables
+all_labels <- list(
   total_malaria_episodes_during_pregnancy ~ "Total Malaria Episodes (Mean ± SD)",
@@
 )
+
+dynamic_labels <- all_labels[names(all_labels) %in% valid_outcomes]
+
+# Define a custom function to format p-values
+format_pvalue <- function(pvals) {
+  sapply(pvals, function(x) {
+    if (is.na(x)) {
+      return("NA")  # Handle missing values properly
+    } else if (is.numeric(x)) {
+      if (x < 0.001) {
+        return("< 0.001")  # Report very small values as "< 0.001"
+      } else if (x < 0.01) {
+        return(format(round(x, 3), nsmall = 3))  # Report between 0.001 and 0.01 to three decimal places
+      } else {
+        return(format(round(x, 2), nsmall = 2))  # Report >= 0.01 to two decimal places
+      }
+    } else {
+      return("NA")  # Return "NA" for non-numeric values
+    }
+  })
+}
+
+# Generate table with formatted p-values
+if (nrow(PROMO_Data) > 0 && length(valid_outcomes) > 1) {
+
+  malaria_outcomes_table <- PROMO_Data %>%
+    select(all_of(valid_outcomes)) %>%
+    tbl_summary(
+      by = "study_arm",
+      statistic = list(
+        total_malaria_episodes_during_pregnancy ~ "{mean} ± {sd}",
+        all_categorical() ~ "{n} ({p}%)"
+      ),
+      label = dynamic_labels
     ) %>%
-  add_p() %>%
-  bold_labels()
+    add_p(
+      test = list(
+        all_categorical() ~ "fisher.test",
+        all_continuous() ~ "t.test"
+      ),
+      pvalue_fun = format_pvalue  # Apply the fixed p-value formatting function
+    )
 
-# Print the formatted table
-malaria_outcomes_table
+  # Print the updated table
+  print(malaria_outcomes_table)
 
-```
-
-*Categorical Characteristics by Study Arm*
-
-```{r}
-# Generate a summary table for categorical characteristics
-categorical_table <- PROMO_Data %>%
-  select(study_arm, stillbirth, neonatal_death, infant_low_birth_weight_2500_g) %>%
-  tbl_summary(by = study_arm, 
-              statistic = list(all_categorical() ~ "{n} ({p}%)"),
-              label = list(stillbirth ~ "Stillbirth",
-                           neonatal_death ~ "Neonatal Death",
-                           infant_low_birth_weight_2500_g ~ "Low Birth Weight")) %>%
-  add_p() %>%
-  bold_labels()
-
-categorical_table
-
+} else {
+  message("⚠️ No data available for analysis or insufficient variability in variables.")
+}
 ```
@@
 ggplot(PROMO_Data, aes(x = total_malaria_episodes_during_pregnancy)) +
-  geom_histogram(binwidth = 1, fill = "blue", color = "black", alpha = 0.7) +
-  labs(title = "Distribution of Malaria Episodes", x = "Malaria Episodes", y = "Count") +
+  geom_histogram(binwidth = 1, fill = "light blue", color = "black", alpha = 0.7) +
+  labs(title = "Distribution of Malaria Episodes During Pregnancy", 
+       x = "Number of Malaria Episodes", 
+       y = "Frequency") +
+  theme_minimal()
+
+```
+
+*Histogram of Gestational Age at Delivery*
+```{r}
+ggplot(PROMO_Data, aes(x = gestational_age_at_delivery_weeks)) +
+  geom_histogram(binwidth = 1, fill = "light blue", color = "black", alpha = 0.7) +
+  labs(title = "Distribution of Gestational Age at Delivery", 
+       x = "Gestational Age (weeks)", 
+       y = "Number of Births") +
+  theme_minimal()
+
+```
+
+*Histogram of Infant Birth Weight*
+```{r}
+ggplot(PROMO_Data, aes(x = infant_birth_weight_kg)) +
+  geom_histogram(binwidth = 0.2, fill = "green", color = "black", alpha = 0.7) +
+  labs(title = "Distribution of Infant Birth Weights", 
+       x = "Birth Weight (kg)", 
+       y = "Number of Infants") +
+  theme_minimal()
+
+```
+
+*Histogram of Maternal Age*
+```{r}
+ggplot(PROMO_Data, aes(x = age_at_enrollment_years)) +
+  geom_histogram(binwidth = 2, fill = "light blue", color = "black", alpha = 0.7) +
+  labs(title = "Distribution of Maternal Age at Enrollment", 
+       x = "Maternal Age (years)", 
+       y = "Number of Participants") +
+  theme_minimal()
+
+```
+
+*Histogram of Malaria Episodes Stratified by Treatment Regimen (SP vs DP)*
+```{r}
+ggplot(PROMO_Data, aes(x = total_malaria_episodes_during_pregnancy, fill = study_arm)) +
+  geom_histogram(binwidth = 1, color = "black", alpha = 0.7, position = "identity") +
+  facet_wrap(~ study_arm) +
+  labs(title = "Malaria Episodes Stratified by Treatment Regimen", 
+       x = "Number of Malaria Episodes", 
+       y = "Frequency") +
   theme_minimal()

### 2025-03-12 products/manuscript/Manuscript.qmd
Project Update

@@
 ---
-knitr:
-  opts_chunk:
-    echo: false
-    message: false
-    warning: false
-
 title: "The Impact of Malaria Episodes and Treatment Regimens on Adverse Pregnancy Outcomes in Ugandan Women"
@@
 csl: ../../assets/american-journal-of-epidemiology.csl
+knitr:
+  opts_chunk:
+    echo: false
+    message: false
+    warning: false
 ---
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
 # Load Required Libraries
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
 # Define the correct file path using the `here` package
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
 # Data Import & Initial Inspection
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
 #  Data Cleaning & Transformation
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
 #  Load required libraries
@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
 # Load required libraries
@@
 
-# Define key variables (check if they exist in dataset)
-malaria_outcome_vars <- c("study_arm", 
+# Define malaria outcome variables
+malaria_outcome_vars <- c(
   "total_malaria_episodes_during_pregnancy",
@@
   "stillbirth",
-                           "infant_low_birth_weight_2500_g")
+  "infant_low_birth_weight_2500_g"
+)
 
-#  Keep only variables that exist in dataset
-malaria_outcome_vars <- intersect(malaria_outcome_vars, colnames(PROMO_Data))
+# Ensure that `study_arm` is in the dataset
+if (!"study_arm" %in% colnames(PROMO_Data)) {
+  stop("⚠️ Error: `study_arm` column not found in PROMO_Data. Please check column names.")
+}
 
-#  Generate the table only if dataset has rows
-if (nrow(PROMO_Data) > 0) {
-  malaria_outcomes_table <- PROMO_Data %>%
-    select(all_of(malaria_outcome_vars)) %>%
-    tbl_summary(
-      by = study_arm,  # Compare DP vs. SP groups
-      statistic = list(
-        total_malaria_episodes_during_pregnancy ~ "{mean} ± {sd}",  # Show mean ± SD
-        all_categorical() ~ "{n} ({p}%)"  # Show counts and percentages for categorical variables
-      ),
-      label = list(
+# Remove `stillbirth` and any other variable that has only one level
+valid_outcomes <- malaria_outcome_vars[sapply(PROMO_Data[malaria_outcome_vars], function(x) length(unique(na.omit(x))) > 1)]
+
+# Ensure `study_arm` is included in the selected variables
+valid_outcomes <- c(valid_outcomes, "study_arm")
+
+# Define label mappings dynamically based on valid variables
+all_labels <- list(
   total_malaria_episodes_during_pregnancy ~ "Total Malaria Episodes (Mean ± SD)",
@@
 )
+
+dynamic_labels <- all_labels[names(all_labels) %in% valid_outcomes]
+
+# Define a custom function to format p-values
+format_pvalue <- function(pvals) {
+  sapply(pvals, function(x) {
+    if (is.na(x)) {
+      return("NA")  # Handle missing values properly
+    } else if (is.numeric(x)) {
+      if (x < 0.001) {
+        return("< 0.001")  # Report very small values as "< 0.001"
+      } else if (x < 0.01) {
+        return(format(round(x, 3), nsmall = 3))  # Report between 0.001 and 0.01 to three decimal places
+      } else {
+        return(format(round(x, 2), nsmall = 2))  # Report >= 0.01 to two decimal places
+      }
+    } else {
+      return("NA")  # Return "NA" for non-numeric values
+    }
+  })
+}
+
+# Generate table with formatted p-values
+if (nrow(PROMO_Data) > 0 && length(valid_outcomes) > 1) {
+
+  malaria_outcomes_table <- PROMO_Data %>%
+    select(all_of(valid_outcomes)) %>%
+    tbl_summary(
+      by = "study_arm",
+      statistic = list(
+        total_malaria_episodes_during_pregnancy ~ "{mean} ± {sd}",
+        all_categorical() ~ "{n} ({p}%)"
+      ),
+      label = dynamic_labels
     ) %>%
-    add_p() %>%  # Add p-values for comparisons
-    modify_header(label = "**Variable**") %>%  # Ensure readable column headers
-    bold_labels()  # Bold table labels for clarity
+    add_p(
+      test = list(
+        all_categorical() ~ "fisher.test",
+        all_continuous() ~ "t.test"
+      ),
+      pvalue_fun = format_pvalue  # Apply the fixed p-value formatting function
+    )
 
-  #  Display only the formatted table
-  malaria_outcomes_table
+  # Print the updated table
+  print(malaria_outcomes_table)
+
 } else {
-  cat("️ No data available for analysis.")
+  message("⚠️ No data available for analysis or insufficient variability in variables.")
 }

### 2025-03-13 products/manuscript/Manuscript.qmd
Project Update

@@
 
-#  Verify Changes
-glimpse(PROMO_Data)  # More readable than str()
-
-#  Summary of Numeric Variables
-print(skim(PROMO_Data))  # Explicitly print skim output
-
-#  Frequency Counts of Key Categorical Variables
-PROMO_Data %>% count(study_arm, sort = TRUE) %>% print()
-PROMO_Data %>% count(placental_malaria, sort = TRUE) %>% print()
-PROMO_Data %>% count(preterm_birth_ontoneo, sort = TRUE) %>% print()
+# Frequency counts of key categorical variables
+# Store counts in variables without printing
+study_arm_counts <- PROMO_Data %>% count(study_arm, sort = TRUE)
+placental_malaria_counts <- PROMO_Data %>% count(placental_malaria, sort = TRUE)
+preterm_birth_counts <- PROMO_Data %>% count(preterm_birth_ontoneo, sort = TRUE)
 
@@
 # Generate table only if dataset has rows
-if (nrow(PROMO_Data) > 0) {
+if (nrow(PROMO_Data) > 0 && length(baseline_vars) > 1) {
+
   baseline_characteristics <- PROMO_Data %>%
     select(all_of(baseline_vars)) %>%
-    tbl_summary(by = study_arm) %>%
+    tbl_summary(
+      by = study_arm,
+      statistic = list(all_categorical() ~ "{n} ({p}%)", all_continuous() ~ "{mean} ± {sd}")
+    ) %>%
     add_p() %>%
@@
 
-  # Display only the formatted table
+  # REMOVE print() → Directly output the table
   baseline_characteristics
+
 } else {
-  cat("️ No data available for analysis.")
+  cat("⚠️ No data available for analysis.")
 }
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
+
 # Load required libraries
@@
 
-  # Print the updated table
-  print(malaria_outcomes_table)
+  # REMOVE print() → Just place the table here
+  malaria_outcomes_table
 
 } else {
-  message("⚠️ No data available for analysis or insufficient variability in variables.")
+  cat("⚠️ No data available for analysis or insufficient variability in variables.")
 }

### 2025-03-14 code/eda-code/eda_MADAprojectcode.qmd
P Update

@@
 ```{r}
-# Read the dataset
+# Data Import & Initial Inspection 
+# load the Dataset
 PROMO_Data <- read_csv(here("data", "raw-data", "PROMO_Data.csv"))
-
-# **Dataset Overview**
-cat("Number of Observations:", nrow(PROMO_Data), "\n")
-cat("Number of Variables:", ncol(PROMO_Data), "\n")
-
-# **Inspect Data Structure**
-str(PROMO_Data)
-head(PROMO_Data)
-
-# **Check Missing Values**
-missing_values <- colSums(is.na(PROMO_Data))
-print(missing_values)
-
-# **Summary Statistics**
-summary(PROMO_Data)
 
@@
 ```{r}
-# Clean Column Names**
-PROMO_Data <- PROMO_Data %>% clean_names()
+#  Clean Column Names
+PROMO_Data <- PROMO_Data %>%
+  clean_names()  # Standardizes column names (lowercase, removes spaces/special characters)
 
-# **Remove Empty Rows**
+#   Check for Missing Values
+missing_summary <- colSums(is.na(PROMO_Data))
+print("Missing Values Summary:")
+print(missing_summary)
+
+#  Handle Missing Values
 PROMO_Data <- PROMO_Data %>%
-  filter(!is.na(study_arm), !is.na(stillbirth), !is.na(neonatal_death), !is.na(infant_low_birth_weight_2500_g))
+  mutate(across(where(is.numeric), ~ ifelse(is.na(.), median(., na.rm = TRUE), .))) %>%  # Replace NA in numeric columns with median
+  mutate(across(where(is.character), ~ replace_na(., "Unknown")))  # Replace NA in categorical columns with "Unknown"
 
-# **Handle Missing Values**
+#   Convert Data Types
 PROMO_Data <- PROMO_Data %>%
-  mutate(across(where(is.numeric), ~ ifelse(is.na(.), median(., na.rm = TRUE), .))) %>%
-  mutate(across(where(is.character), ~ replace_na(., "Unknown")))
+  mutate(across(c(study_arm, delivery_mode, education_level), as.factor)) %>% # Convert categorical columns to factors
+  mutate(across(contains("date"), as.Date, format="%Y-%m-%d"))  # Convert date columns
 
-# **Convert Data Types**
+#  Rename Key Variables for Clarity
 PROMO_Data <- PROMO_Data %>%
-  mutate(across(c(study_arm, stillbirth, neonatal_death, placental_malaria, delivery_mode), as.factor)) %>%
-  mutate(across(contains("date"), as.Date, format="%Y-%m-%d"))
+  rename(
+    `Study Arm` = study_arm,
+    `Gestational Age at Enrollment (Weeks)` = gestational_age_at_enrollment_weeks,
+    `Gestational Age at Delivery (Weeks)` = gestational_age_at_delivery,
+    `Education Level` = education_level,
+    `Detection of Malaria Parasites by Microscopy` = proportion_of_plasmodium_positive_blood_smears_during_pregnancy_by_microscopy,
+    `Detection of Malaria Parasites by Microscopy or qPCR` = total_plasmodium_positive_routine_blood_smears_during_pregnancy
+  )
 
-# **Verify Changes**
-str(PROMO_Data)
+#   Remove Duplicate Rows
+PROMO_Data_Cleaned <- PROMO_Data %>%
+  distinct()  # Removes exact duplicate rows
+
+#  Step 7: Print Summary of Cleaned Data
+summary(PROMO_Data_Cleaned)
+
+#   Display Column Names to Verify
+print("Updated Column Names:")
+print(colnames(PROMO_Data_Cleaned))
+
+```
+
+```{r}
+colnames(PROMO_Data_Cleaned)
 
@@
 # Exploratory Data Analysis (EDA)
+```{r}
+# Number of observations and variables
+cat("Number of Observations:", nrow(PROMO_Data_Cleaned), "\n")
+cat("Number of Variables:", ncol(PROMO_Data_Cleaned), "\n")
 
-```{r}
-# **Summary of Numeric Variables**
-skim(PROMO_Data)
+# View structure of dataset
+str(PROMO_Data_Cleaned)
 
-# **Frequency Counts of Key Categorical Variables**
-PROMO_Data %>%
-  count(study_arm, sort = TRUE)
-
-PROMO_Data %>%
-  count(placental_malaria, sort = TRUE)
-
-PROMO_Data %>%
-  count(preterm_birth_ontoneo, sort = TRUE)
+# Quick Summary Statistics
+summary(PROMO_Data_Cleaned)
 
@@
 
-# Characteristics Tables
-*Baseline Characteristics Table*
+*Step 1: Fix Missing Data & Categorize Variables*
+*Baseline Characteristics Table 1*
+```{r}
+library(dplyr)
+library(gtsummary)
+
+# ✅ Fix & Categorize Variables
+PROMO_Data_Cleaned <- PROMO_Data_Cleaned %>%
+  mutate(
+    Age_Category = case_when(
+      age_at_enrollment_years < 18 ~ "<18 (Minors)",
+      age_at_enrollment_years >= 18 & age_at_enrollment_years <= 35 ~ "18–35 (Young Adults)",
+      age_at_enrollment_years >= 36 & age_at_enrollment_years <= 55 ~ "36–55 (Middle-aged)",
+      age_at_enrollment_years > 55 ~ ">55 (Older Adults)",
+      TRUE ~ "Unknown"
+    ),
+
+    # ✅ Fix Education Level Categorization (Put "Unknown" Last)
+    `Education Level` = case_when(
+      `Education Level` %in% c("A level", "O level", "None") ~ "Secondary/No Formal Education",
+      `Education Level` == "Primary" ~ "Primary",
+      `Education Level` %in% c("Tertiary", "University") ~ "Tertiary/Higher Education",
+      is.na(`Education Level`) ~ "Unknown",
+      TRUE ~ "Unknown"
+    ) %>% factor(levels = c("Primary", "Secondary/No Formal Education", "Tertiary/Higher Education", "Unknown")),
+
+    # ✅ Fix Gravidity Categorization (Put "Unknown" Last)
+    Gravidity = case_when(
+      gravidity == 1 ~ "1",
+      gravidity == 2 ~ "2",
+      gravidity >= 3 ~ "≥3",
+      is.na(gravidity) ~ "Unknown",
+      TRUE ~ "Unknown"
+    ) %>% factor(levels = c("1", "2", "≥3", "Unknown")),
+
+    # ✅ Fix Parity Categorization (Put "Unknown" Last)
+    Parity = case_when(
+      parity == 1 ~ "1",
+      parity == 2 ~ "2",
+      parity >= 3 ~ "≥3",
+      is.na(parity) ~ "Unknown",
+      TRUE ~ "Unknown"
+    ) %>% factor(levels = c("1", "2", "≥3", "Unknown"))
+  )
+# ✅ Define Variables (Maternal Info First, Infant Last)
+baseline_vars <- c(
+  # **Maternal Information**
+  "Age_Category", 
+  "Education Level",
+  "Gravidity", 
+  "Parity", 
+  "Gestational Age at Enrollment (Weeks)", 
+  "Gestational Age at Delivery (Weeks)", 
+  "placental_malaria", 
+  "total_malaria_episodes_during_pregnancy", 
+  "delivery_mode",
+
+  # **Pregnancy Outcomes**
+  "preterm_birth", 
+  "small_for_gestational_age", 
+  "spontaneous_abortion", 
+  "stillbirth", 
+  "neonatal_death",
+
+  # **Infant Information**
+  "infant_low_birth_weight_2500_g", 
+  "infant_sex6"
+)
+
+# ✅ Ensure Variables Exist in the Dataset
+baseline_vars <- baseline_vars[baseline_vars %in% colnames(PROMO_Data_Cleaned)]
+
+# ✅ Create Updated Table (Remove Label Name)
+baseline_table <- PROMO_Data_Cleaned %>%
+  select(all_of(baseline_vars), `Study Arm`) %>%
+  tbl_summary(
+    by = `Study Arm`,
+    statistic = all_categorical() ~ "{n} ({p}%)",
+    missing = "no"
+  ) %>%
+  modify_header(label = "") %>% # Remove the label name
+  bold_labels()
+
+# ✅ Display Table
+baseline_table
+```
 
 ```{r}
-# Select key variables for baseline characteristics
-baseline_vars <- c("study_arm", "education_level_omrse", "gravidity", "parity", 
-                   "placental_malaria", "preterm_birth_ontoneo", "stillbirth", 
-                   "neonatal_death", "infant_low_birth_weight_2500_g")
+PROMO_Data_Cleaned <- PROMO_Data_Cleaned %>%
+  mutate(
+    Age_Category = case_when(
+      age_at_enrollment_years < 18 ~ "<18 (Minors)",
+      age_at_enrollment_years >= 18 & age_at_enrollment_years <= 35 ~ "18–35 (Young Adults)",
+      age_at_enrollment_years >= 36 & age_at_enrollment_years <= 55 ~ "36–55 (Middle-aged)",
+      age_at_enrollment_years > 55 ~ ">55 (Older Adults)",
+      TRUE ~ "Unknown"
+    ),
 
-# **Ensure selected variables exist**
-baseline_vars <- baseline_vars[baseline_vars %in% colnames(PROMO_Data)]
+    # ✅ Fix Education Level Categorization (Include "Unknown")
+    `Education Level` = case_when(
+      `Education Level` %in% c("A level", "O level", "None") ~ "Secondary/No Formal Education",
+      `Education Level` == "Primary" ~ "Primary",
+      `Education Level` %in% c("Tertiary", "University") ~ "Tertiary/Higher Education",
+      is.na(`Education Level`) ~ "Unknown",
+      TRUE ~ "Unknown"
+    ) %>% factor(levels = c("Primary", "Secondary/No Formal Education", "Tertiary/Higher Education", "Unknown")),
 
-# Generate formatted table using gtsummary
-baseline_characteristics <- PROMO_Data %>%
-  select(all_of(baseline_vars)) %>%
-  tbl_summary(by = study_arm) %>%
-  add_p() %>%
-  modify_header(label = "**Variable**") %>%
+    # ✅ Fix Gravidity Categorization (Include "Unknown")
+    gravidity = case_when(
+      gravidity == 1 ~ "1",
+      gravidity == 2 ~ "2",
+      gravidity >= 3 ~ "≥3",
+      is.na(gravidity) ~ "Unknown",
+      TRUE ~ "Unknown"
+    ) %>% factor(levels = c("1", "2", "≥3", "Unknown")),
+
+    # ✅ Fix Parity Categorization (Include "Unknown")
+    parity = case_when(
+      parity == 1 ~ "1",
+      parity == 2 ~ "2",
+      parity >= 3 ~ "≥3",
+      is.na(parity) ~ "Unknown",
+      TRUE ~ "Unknown"
+    ) %>% factor(levels = c("1", "2", "≥3", "Unknown"))
+  )
+
+PROMO_Data_Cleaned <- PROMO_Data_Cleaned %>%
+  mutate(
+    Age_Category = case_when(
+      age_at_enrollment_years < 18 ~ "<18 (Minors)",
+      age_at_enrollment_years >= 18 & age_at_enrollment_years <= 35 ~ "18–35 (Young Adults)",
+      age_at_enrollment_years >= 36 & age_at_enrollment_years <= 55 ~ "36–55 (Middle-aged)",
+      age_at_enrollment_years > 55 ~ ">55 (Older Adults)",
+      TRUE ~ "Unknown"
+    ),
+
+    # ✅ Fix Education Level Categorization (Include "Unknown")
+    `Education Level` = case_when(
+      `Education Level` %in% c("A level", "O level", "None") ~ "Secondary/No Formal Education",
+      `Education Level` == "Primary" ~ "Primary",
+      `Education Level` %in% c("Tertiary", "University") ~ "Tertiary/Higher Education",
+      is.na(`Education Level`) ~ "Unknown",
+      TRUE ~ "Unknown"
+    ) %>% factor(levels = c("Primary", "Secondary/No Formal Education", "Tertiary/Higher Education", "Unknown")),
+
+    # ✅ Fix Gravidity Categorization (Include "Unknown")
+    gravidity = case_when(
+      gravidity == 1 ~ "1",
+      gravidity == 2 ~ "2",
+      gravidity >= 3 ~ "≥3",
+      is.na(gravidity) ~ "Unknown",
+      TRUE ~ "Unknown"
+    ) %>% factor(levels = c("1", "2", "≥3", "Unknown")),
+
+    # ✅ Fix Parity Categorization (Include "Unknown")
+    parity = case_when(
+      parity == 1 ~ "1",
+      parity == 2 ~ "2",
+      parity >= 3 ~ "≥3",
+      is.na(parity) ~ "Unknown",
+      TRUE ~ "Unknown"
+    ) %>% factor(levels = c("1", "2", "≥3", "Unknown"))
+  )
+
+# ✅ Define Baseline Characteristics Variables
+baseline_vars <- c(
+  # **Maternal Information First**
+  "Age_Category", 
+  "Education Level",
+  "Gravidity", 
+  "Parity", 
+  "Gestational Age at Enrollment (Weeks)", 
+  "Gestational Age at Delivery (Weeks)", 
+  "placental_malaria", 
+  "total_malaria_episodes_during_pregnancy", 
+  "delivery_mode",
+
+  # **Pregnancy Outcomes**
+  "preterm_birth", 
+  "small_for_gestational_age", 
+  "spontaneous_abortion", 
+  "stillbirth", 
+  "neonatal_death",
+
+  # **Infant Information**
+  "infant_low_birth_weight_2500_g", 
+  "infant_sex6"
+)
+
+# ✅ Ensure Variables Exist in the Dataset
+baseline_vars <- baseline_vars[baseline_vars %in% colnames(PROMO_Data_Cleaned)]
+
+# ✅ Create Updated Baseline Characteristics Table
+baseline_table <- PROMO_Data_Cleaned %>%
+  select(all_of(baseline_vars), `Study Arm`) %>%
+  tbl_summary(
+    by = `Study Arm`,
+    statistic = all_categorical() ~ "{n} ({p}%)",
+    missing = "no"
+  ) %>%
+  modify_header(label = "**Baseline Characteristics**") %>%
   bold_labels()
 
-baseline_characteristics
-
-```
-
-*Continuous Characteristics by Study Arm*
-
-```{r}
-# Function to summarize continuous variables
-summarize_continuous <- function(var) {
-  sp_mean <- mean(PROMO_Data[[var]][PROMO_Data$study_arm == "Sulphadoxine-pyrimethamine (SP)"], na.rm = TRUE)
-  sp_sd <- sd(PROMO_Data[[var]][PROMO_Data$study_arm == "Sulphadoxine-pyrimethamine (SP)"], na.rm = TRUE)
-
-  dp_mean <- mean(PROMO_Data[[var]][PROMO_Data$study_arm == "Dihydroartemisinin-piperaquine (DP)"], na.rm = TRUE)
-  dp_sd <- sd(PROMO_Data[[var]][PROMO_Data$study_arm == "Dihydroartemisinin-piperaquine (DP)"], na.rm = TRUE)
-
-  p_value <- t.test(PROMO_Data[[var]] ~ PROMO_Data$study_arm, var.equal = TRUE, na.rm = TRUE)$p.value
-
-  return(c(sprintf("%.1f (%.1f)", sp_mean, sp_sd),
-           sprintf("%.1f (%.1f)", dp_mean, dp_sd),
-           sprintf("%.3f", p_value)))
-}
-
-# Create summary table
-summary_table <- tibble(
-  Characteristic = c("Maternal Age (years)", "Gravidity", "Gestational Age at Delivery (weeks)", "Malaria Episodes During Pregnancy"),
-  `SP (Mean ± SD)` = sapply(c("age_at_enrollment_years", "gravidity", "gestational_age_at_delivery_weeks", "total_malaria_episodes_during_pregnancy"), summarize_continuous)[1,],
-  `DP (Mean ± SD)` = sapply(c("age_at_enrollment_years", "gravidity", "gestational_age_at_delivery_weeks", "total_malaria_episodes_during_pregnancy"), summarize_continuous)[2,],
-  `p-value` = sapply(c("age_at_enrollment_years", "gravidity", "gestational_age_at_delivery_weeks", "total_malaria_episodes_during_pregnancy"), summarize_continuous)[3,]
-)
-
-# Display table using gt
-summary_table %>%
-  gt() %>%
-  tab_header(title = md("**Table: Continuous Characteristics by Study Arm**")) %>%
-  cols_align(align = "center") %>%
-  tab_options(
-    table.font.size = px(12),
-    heading.title.font.size = px(14),
-    table.border.top.style = "solid",
-    table.border.bottom.style = "solid"
-  )
+# ✅ Display the Table
+baseline_table
 
@@
 ```{r}
-colnames(PROMO_Data)
+# Load required packages
+library(dplyr)
+library(gtsummary)
+
+# ✅ Step 1: Categorize Age into Clinically Meaningful Groups
+PROMO_Data_Cleaned <- PROMO_Data_Cleaned %>%
+  mutate(
+    # ✅ Age Category
+    Age_Category = case_when(
+      age_at_enrollment_years < 18 ~ "<18 (Minors)",
+      age_at_enrollment_years >= 18 & age_at_enrollment_years <= 35 ~ "18–35 (Young Adults)",
+      age_at_enrollment_years >= 36 & age_at_enrollment_years <= 55 ~ "36–55 (Middle-aged)",
+      age_at_enrollment_years > 55 ~ ">55 (Older Adults)",
+      TRUE ~ NA_character_
+    ),
+
+    # ✅ Fix Education Level Categorization (Ensuring it’s Correct)
+    Education_Level = case_when(
+      `Education Level` %in% c("A level", "O level", "None") ~ "Secondary/No Formal Education",
+      `Education Level` %in% c("Primary") ~ "Primary",
+      `Education Level` %in% c("Tertiary", "University") ~ "Tertiary/Higher Education",
+      TRUE ~ NA_character_
+    ),
+
+    # Convert to Factor with Proper Levels
+    Education_Level = factor(Education_Level, levels = c("Primary", "Secondary/No Formal Education", "Tertiary/Higher Education")),
+
+    # ✅ Categorize Gravidity
+    Gravidity = factor(case_when(
+      gravidity == 1 ~ "1",
+      gravidity == 2 ~ "2",
+      gravidity >= 3 ~ "≥3",
+      TRUE ~ NA_character_
+    ), levels = c("1", "2", "≥3")),
+
+    # ✅ Categorize Parity
+    Parity = factor(case_when(
+      parity == 1 ~ "1",
+      parity == 2 ~ "2",
+      parity >= 3 ~ "≥3",
+      TRUE ~ NA_character_
+    ), levels = c("1", "2", "≥3"))
+  )
+
+# ✅ Step 2: Define the 16 Key Variables for the Table
+baseline_vars <- c(
+  # 🟢 **Mother's Demographics & Reproductive History**
+  "Age_Category", 
+  "Education_Level",  # ✅ Fixed Education Level Here
+  "Gravidity",
+  "Parity",
+  "Gestational Age at Enrollment (Weeks)", 
+  "Gestational Age at Delivery (Weeks)", 
+  "Placental_malaria", 
+  "Total_malaria_episodes_during_pregnancy",
+  "Study Arm",  # Treatment assignment (SP vs DP)
+  "Delivery_mode",
+
+  # 🟡 **Pregnancy & Birth Outcomes**
+  "Preterm_birth", 
+  "Small_for_gestational_age", 
+  "Spontaneous_abortion", 
+  "Stillbirth", 
+  "Neonatal_death",
+
+  # 🔵 **Infant Birth Outcomes**
+  "Infant_low_birth_weight_2500_g", 
+  "Infant_sex6"
+)
+
+# ✅ Step 3: Ensure Variables Exist in the Dataset
+baseline_vars <- baseline_vars[baseline_vars %in% colnames(PROMO_Data_Cleaned)]
+
+# ✅ Step 4: Confirm Total Variables
+cat("Total Variables Selected:", length(baseline_vars), "\n")  # Should output 16
+
+# ✅ Step 5: Create the Baseline Characteristics Table
+baseline_table <- PROMO_Data_Cleaned %>%
+  select(all_of(baseline_vars), `Study Arm`) %>%
+  tbl_summary(
+    by = `Study Arm`,
+    statistic = all_categorical() ~ "{n} ({p}%)",
+    missing = "no"
+  ) %>%
+  modify_header(label = "**Baseline Characteristics**") %>%
+  bold_labels()
+
+# ✅ Step 6: Display the Table
+baseline_table
+
+```
+
+# Data Visualization
+
+*Figure 1: Pregnancy Outcomes by Study Arm*
+```{r}
+# Load required packages
+library(ggplot2)
+library(dplyr)
+library(tidyr)
+
+# Ensure correct column name for Study Arm
+study_arm_col <- "Study Arm"
+
+# Define pregnancy outcomes
+pregnancy_outcomes <- c("preterm_birth", "stillbirth", "neonatal_death")
+
+# Reshape data for plotting
+pregnancy_data <- PROMO_Data_Cleaned %>%
+  select(all_of(study_arm_col), all_of(pregnancy_outcomes)) %>%
+  pivot_longer(cols = all_of(pregnancy_outcomes), names_to = "Outcome", values_to = "Present") %>%
+  filter(!is.na(Present)) %>%  # Remove missing values
+  group_by(across(all_of(study_arm_col)), Outcome, Present) %>%
+  summarise(Total = n(), .groups = "drop")
+
+# Convert outcome values to categorical (if numeric)
+pregnancy_data$Present <- factor(pregnancy_data$Present, levels = c(0, 1), labels = c("Absent", "Present"))
+
+# Plot Pregnancy Outcomes by Study Arm
+ggplot(pregnancy_data, aes(x = Outcome, y = Total, fill = !!sym(study_arm_col))) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Pregnancy Outcomes by Study Arm",
+       x = "Pregnancy Outcome",
+       y = "Total Count",
+       fill = "Study Arm") +
+  theme_minimal() +
+  scale_fill_manual(values = c("lightblue", "pink")) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+```
+
+*Placental Malaria by Study Arm*
+```{r}
+# Group and summarize data for placental malaria
+# Ensure correct column name for Study Arm
+study_arm_col <- "Study Arm"
+
+# Group and summarize data for placental malaria
+placental_malaria_data <- PROMO_Data_Cleaned %>%
+  group_by(across(all_of(study_arm_col)), placental_malaria) %>%
+  summarise(Count = n(), .groups = "drop")
+
+# Convert placental malaria values to categorical (if numeric)
+placental_malaria_data$placental_malaria <- factor(placental_malaria_data$placental_malaria, 
+                                                   levels = c(0, 1), 
+                                                   labels = c("No", "Yes"))
+
+# Plot Placental Malaria by Study Arm
+ggplot(placental_malaria_data, aes(x = placental_malaria, y = Count, fill = !!sym(study_arm_col))) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Placental Malaria by Study Arm",
+       x = "Placental Malaria",
+       y = "Total Count",
+       fill = "Study Arm") +
+  theme_minimal() +
+  scale_fill_manual(values = c("lightblue", "lightgreen")) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
@@
 # Load required libraries
-# Load required libraries
 library(dplyr)
 library(gtsummary)
-
-# Define malaria outcome variables
-malaria_outcome_vars <- c(
-  "total_malaria_episodes_during_pregnancy",
-  "placental_malaria",
-  "preterm_birth_ontoneo",
-  "stillbirth",
-  "infant_low_birth_weight_2500_g"
-)
-
-# Ensure that `study_arm` is in the dataset
-if (!"study_arm" %in% colnames(PROMO_Data)) {
-  stop("⚠️ Error: `study_arm` column not found in PROMO_Data. Please check column names.")
-}
-
-# Remove `stillbirth` and any other variable that has only one level
-valid_outcomes <- malaria_outcome_vars[sapply(PROMO_Data[malaria_outcome_vars], function(x) length(unique(na.omit(x))) > 1)]
-
-# Ensure `study_arm` is included in the selected variables
-valid_outcomes <- c(valid_outcomes, "study_arm")
-
-# Define label mappings dynamically based on valid variables
-all_labels <- list(
-  total_malaria_episodes_during_pregnancy ~ "Total Malaria Episodes (Mean ± SD)",
-  placental_malaria ~ "Placental Malaria (Histology)",
-  preterm_birth_ontoneo ~ "Preterm Birth",
-  stillbirth ~ "Stillbirth",
-  infant_low_birth_weight_2500_g ~ "Low Birth Weight (<2500g)"
-)
-
-dynamic_labels <- all_labels[names(all_labels) %in% valid_outcomes]
-
-# Define a custom function to format p-values
-format_pvalue <- function(pvals) {
-  sapply(pvals, function(x) {
-    if (is.na(x)) {
-      return("NA")  # Handle missing values properly
-    } else if (is.numeric(x)) {
-      if (x < 0.001) {
-        return("< 0.001")  # Report very small values as "< 0.001"
-      } else if (x < 0.01) {
-        return(format(round(x, 3), nsmall = 3))  # Report between 0.001 and 0.01 to three decimal places
-      } else {
-        return(format(round(x, 2), nsmall = 2))  # Report >= 0.01 to two decimal places
-      }
-    } else {
-      return("NA")  # Return "NA" for non-numeric values
-    }
-  })
-}
-
-# Generate table with formatted p-values
-if (nrow(PROMO_Data) > 0 && length(valid_outcomes) > 1) {
-
-  malaria_outcomes_table <- PROMO_Data %>%
-    select(all_of(valid_outcomes)) %>%
-    tbl_summary(
-      by = "study_arm",
-      statistic = list(
-        total_malaria_episodes_during_pregnancy ~ "{mean} ± {sd}",
-        all_categorical() ~ "{n} ({p}%)"
-      ),
-      label = dynamic_labels
-    ) %>%
-    add_p(
-      test = list(
-        all_categorical() ~ "fisher.test",
-        all_continuous() ~ "t.test"
-      ),
-      pvalue_fun = format_pvalue  # Apply the fixed p-value formatting function
-    )
-
-  # Print the updated table
-  print(malaria_outcomes_table)
-
-} else {
-  message("⚠️ No data available for analysis or insufficient variability in variables.")
-}
-```
 

### 2025-03-14 products/manuscript/Manuscript.qmd
P Update

--- 2025-03-12 products/manuscript/Manuscript.qmd
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-# Define the correct file path using the `here` package
-data_path <- here("data", "raw-data", "PROMO_Data.csv")
 
-# Read the CSV file, suppressing messages
-PROMO_Data <- read_csv(data_path, show_col_types = FALSE)
+# Load necessary packages
+library(dplyr)
+library(readr)
+library(janitor)
+library(tidyr)
+library(here)
+library(knitr)  # For table output
+
+# ✅ Define the correct file path using `here` package
+PROMO_Data <- read_csv(here("data", "raw-data", "PROMO_Data.csv"), show_col_types = FALSE)
+
+# ✅ Clean Column Names
+PROMO_Data <- PROMO_Data %>%
+  clean_names()  # Standardizes column names (lowercase, removes spaces/special characters)
+
+# ✅ Handle Missing Values
+PROMO_Data <- PROMO_Data %>%
+  mutate(across(where(is.numeric), ~ ifelse(is.na(.), median(., na.rm = TRUE), .))) %>%  # Replace NA in numeric columns with median
+  mutate(across(where(is.character), ~ replace_na(., "Unknown")))  # Replace NA in categorical columns with "Unknown"
+
+# ✅ Convert Data Types
+PROMO_Data <- PROMO_Data %>%
+  mutate(across(c(study_arm, delivery_mode, education_level), as.factor)) %>% # Convert categorical columns to factors
+  mutate(across(contains("date"), as.Date, format="%Y-%m-%d"))  # Convert date columns
+
+# ✅ Rename Key Variables for Clarity
+PROMO_Data <- PROMO_Data %>%
+  rename(
+    `Study Arm` = study_arm,
+    `Gestational Age at Enrollment (Weeks)` = gestational_age_at_enrollment_weeks,
+    `Gestational Age at Delivery (Weeks)` = gestational_age_at_delivery,
+    `Education Level` = education_level,
+    `Detection of Malaria Parasites by Microscopy` = proportion_of_plasmodium_positive_blood_smears_during_pregnancy_by_microscopy,
+    `Detection of Malaria Parasites by Microscopy or qPCR` = total_plasmodium_positive_routine_blood_smears_during_pregnancy
+  )
+
+# ✅ Remove Duplicate Rows
+PROMO_Data_Cleaned <- PROMO_Data %>%
+  distinct()  # Removes exact duplicate rows
+
+# ✅ Standardize Categorical Variables with Proper Ordering
+PROMO_Data_Cleaned <- PROMO_Data_Cleaned %>%
+  mutate(
+    # Age Categories
+    Age_Category = factor(case_when(
+      age_at_enrollment_years < 18 ~ "<18",
+      age_at_enrollment_years >= 18 & age_at_enrollment_years <= 35 ~ "18–35",
+      age_at_enrollment_years >= 36 & age_at_enrollment_years <= 55 ~ "36–55",
+      age_at_enrollment_years > 55 ~ ">55",
+      TRUE ~ "Unknown"
+    ), levels = c("<18", "18–35", "36–55", ">55", "Unknown")),
+
+    # Education Level Categories (With "Unknown" at the End)
+    `Education Level` = factor(case_when(
+      `Education Level` %in% c("A level", "O level", "None") ~ "Secondary/No Formal Education",
+      `Education Level` == "Primary" ~ "Primary",
+      `Education Level` %in% c("Tertiary", "University") ~ "Tertiary/Higher Education",
+      is.na(`Education Level`) ~ "Unknown",
+      TRUE ~ "Unknown"
+    ), levels = c("Primary", "Secondary/No Formal Education", "Tertiary/Higher Education", "Unknown")),
+
+    # Gravidity (Number of Pregnancies) with Proper Ordering
+    Gravidity = factor(case_when(
+      gravidity == 1 ~ "1",
+      gravidity == 2 ~ "2",
+      gravidity >= 3 ~ "≥3",
+      is.na(gravidity) ~ "Unknown",
+      TRUE ~ "Unknown"
+    ), levels = c("1", "2", "≥3", "Unknown")),
+
+    # Parity (Number of Births) with Proper Ordering
+    Parity = factor(case_when(
+      parity == 1 ~ "1",
+      parity == 2 ~ "2",
+      parity >= 3 ~ "≥3",
+      is.na(parity) ~ "Unknown",
+      TRUE ~ "Unknown"
+    ), levels = c("1", "2", "≥3", "Unknown"))
+  )
+
+# ✅ Create a Summary Table
+summary_table <- PROMO_Data_Cleaned %>%
+  summarise(
+    `Total Participants` = n(),
+    `Median Age` = median(age_at_enrollment_years, na.rm = TRUE),
+    `Median Gestational Age at Enrollment` = median(`Gestational Age at Enrollment (Weeks)`, na.rm = TRUE),
+    `Median Gestational Age at Delivery` = median(`Gestational Age at Delivery (Weeks)`, na.rm = TRUE)
+  )
+
 ```
@@
 Despite significant efforts in prevention and treatment, malaria remains a major global health challenge.
+
 Vulnerable groups such as pregnant women, children, and immunocompromised individuals are at the highest risk, and Africa continues to bear the heaviest burden of this disease.
@@
 
+Table 1: Baseline Characteristics and Pregnancy Outcomes of Pregnant Women
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-# Data Import & Initial Inspection
 
-# Read the dataset
+# Load Required Packages
+library(dplyr)
+library(gtsummary)
+library(janitor)
+library(here)
+library(knitr)
+
+# ✅ Load Dataset (Ensure Correct File Path)
 PROMO_Data <- read_csv(here("data", "raw-data", "PROMO_Data.csv"), show_col_types = FALSE)
-```
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-#  Data Cleaning & Transformation
+# ✅ Clean Column Names
+PROMO_Data <- PROMO_Data %>%
+  clean_names()  # Standardizes names (lowercase, removes spaces/special characters)
 
-# Load required libraries (if not already loaded)
-library(janitor)   # For clean_names()
-
-# Clean Column Names
-PROMO_Data <- PROMO_Data %>% clean_names()
-
-# Remove Empty Rows
+# ✅ Handle Missing Values & Convert Data Types
 PROMO_Data <- PROMO_Data %>%
-  filter(!is.na(study_arm), !is.na(stillbirth), !is.na(neonatal_death), !is.na(infant_low_birth_weight_2500_g))
-
-#  Handle Missing Values
-PROMO_Data <- PROMO_Data %>%
-  replace_na(list(
-    study_arm = "Unknown",
-    stillbirth = "Unknown",
-    neonatal_death = "Unknown",
-    placental_malaria = "Unknown",
-    delivery_mode = "Unknown"
-  )) %>%
-  mutate(across(where(is.numeric), ~ ifelse(is.na(.), median(., na.rm = TRUE), .)))
-
-#  Convert Data Types
-PROMO_Data <- PROMO_Data %>%
-  mutate(across(c(study_arm, stillbirth, neonatal_death, placental_malaria, delivery_mode), as.factor)) %>%
+  mutate(across(where(is.numeric), ~ ifelse(is.na(.), median(., na.rm = TRUE), .))) %>%
+  mutate(across(where(is.character), ~ replace_na(., "Unknown"))) %>%
+  mutate(across(c(study_arm, delivery_mode, education_level), as.factor)) %>%
   mutate(across(contains("date"), as.Date, format="%Y-%m-%d"))
 
-#  Verify Changes
-glimpse(PROMO_Data)  # More readable than str()
+# ✅ Rename Key Variables for Clarity
+PROMO_Data <- PROMO_Data %>%
+  rename(
+    `Study Arm` = study_arm,
+    `Gestational Age at Enrollment (Weeks)` = gestational_age_at_enrollment_weeks,
+    `Gestational Age at Delivery (Weeks)` = gestational_age_at_delivery,
+    `Education Level` = education_level,
+    `Placental Malaria` = placental_malaria,
+    `Total Malaria Episodes During Pregnancy` = total_malaria_episodes_during_pregnancy
+  )
 
-#  Summary of Numeric Variables
-print(skim(PROMO_Data))  # Explicitly print skim output
+# ✅ Remove Duplicate Rows
+PROMO_Data_Cleaned <- PROMO_Data %>% distinct()
 
-#  Frequency Counts of Key Categorical Variables
-PROMO_Data %>% count(study_arm, sort = TRUE) %>% print()
-PROMO_Data %>% count(placental_malaria, sort = TRUE) %>% print()
-PROMO_Data %>% count(preterm_birth_ontoneo, sort = TRUE) %>% print()
+# ✅ Standardize Categorical Variables (Ensuring "Unknown" Last)
+PROMO_Data_Cleaned <- PROMO_Data_Cleaned %>%
+  mutate(
+    # ✅ Age Category (Ensure Correct Order)
+    Age_Category = factor(case_when(
+      age_at_enrollment_years < 18 ~ "<18",
+      age_at_enrollment_years >= 18 & age_at_enrollment_years <= 35 ~ "18–35",
+      age_at_enrollment_years >= 36 & age_at_enrollment_years <= 55 ~ "36–55",
+      age_at_enrollment_years > 55 ~ ">55 (Older Adults)",
+      TRUE ~ "Unknown"
+    ), levels = c("<18", "18–35", "36–55", ">55 (Older Adults)", "Unknown")),
+
+    # ✅ Education Level (Ensure "Unknown" is Last)
+    `Education Level` = factor(case_when(
+      `Education Level` %in% c("A level", "O level", "None") ~ "Secondary/No Formal Education",
+      `Education Level` == "Primary" ~ "Primary",
+      `Education Level` %in% c("Tertiary", "University") ~ "Tertiary/Higher Education",
+      is.na(`Education Level`) ~ "Unknown",
+      TRUE ~ "Unknown"
+    ), levels = c("Primary", "Secondary/No Formal Education", "Tertiary/Higher Education", "Unknown")),
+
+    # ✅ Gravidity (Ensure "Unknown" is Last)
+    Gravidity = factor(case_when(
+      gravidity == 1 ~ "1",
+      gravidity == 2 ~ "2",
+      gravidity >= 3 ~ "≥3",
+      is.na(gravidity) ~ "Unknown",
+      TRUE ~ "Unknown"
+    ), levels = c("1", "2", "≥3", "Unknown")),
+
+    # ✅ Parity (Ensure "Unknown" is Last)
+    Parity = factor(case_when(
+      parity == 1 ~ "1",
+      parity == 2 ~ "2",
+      parity >= 3 ~ "≥3",
+      is.na(parity) ~ "Unknown",
+      TRUE ~ "Unknown"
+    ), levels = c("1", "2", "≥3", "Unknown"))
+  )
+
+# ✅ Define Baseline Characteristics Variables (Reorganized for Logical Flow)
+baseline_vars <- c(
+  # **Maternal Information First**
+  "Age_Category", 
+  "Education Level",
+  "Gravidity", 
+  "Parity", 
+  "Placental Malaria", 
+  "Total Malaria Episodes During Pregnancy", 
+  "delivery_mode",
+
+  # **Pregnancy Outcomes**
+  "preterm_birth", 
+  "small_for_gestational_age", 
+  "spontaneous_abortion", 
+  "stillbirth", 
+  "neonatal_death",
+
+  # **Infant Information**
+  "infant_low_birth_weight_2500_g", 
+  "infant_sex6"
+)
+
+# ✅ Ensure Selected Variables Exist in Dataset
+baseline_vars <- baseline_vars[baseline_vars %in% colnames(PROMO_Data_Cleaned)]
+
+# ✅ Generate Baseline Characteristics Table
+baseline_table <- PROMO_Data_Cleaned %>%
+  select(all_of(baseline_vars), `Study Arm`) %>%
+  tbl_summary(
+    by = `Study Arm`,
+    statistic = all_categorical() ~ "{n} ({p}%)",
+    missing = "no"
+  ) %>%
+  modify_header(label = "Characteristics") %>%
+  bold_labels()
+
+# ✅ Display the Table for Quarto Word Output
+baseline_table
 
@@
 
+Figure 1. Distribution of Adverse Pregnancy Outcomes by Study Arm
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-#  Load required libraries
+# Load required packages
+library(ggplot2)
 library(dplyr)
-library(gtsummary)
+library(tidyr)
 
-#  Select key baseline variables (ensure they exist)
-baseline_vars <- c("study_arm", "education_level_omrse", "gravidity", "parity", 
-                   "placental_malaria", "preterm_birth_ontoneo", "stillbirth", 
-                   "neonatal_death", "infant_low_birth_weight_2500_g")
+# Ensure correct column name for Study Arm
+study_arm_col <- "Study Arm"
 
-#  Ensure only existing variables are used
-baseline_vars <- intersect(baseline_vars, colnames(PROMO_Data))
+# Define pregnancy outcomes
+pregnancy_outcomes <- c("preterm_birth", "stillbirth", "neonatal_death")
 
-# Generate table only if dataset has rows
-if (nrow(PROMO_Data) > 0) {
-  baseline_characteristics <- PROMO_Data %>%
-    select(all_of(baseline_vars)) %>%
-    tbl_summary(by = study_arm) %>%
-    add_p() %>%
-    modify_header(label = "**Variable**") %>%
-    bold_labels()
+# ✅ Reshape data for plotting
+pregnancy_data <- PROMO_Data_Cleaned %>%
+  select(all_of(study_arm_col), all_of(pregnancy_outcomes)) %>%
+  pivot_longer(cols = all_of(pregnancy_outcomes), names_to = "Outcome", values_to = "Present") %>%
+  filter(!is.na(Present)) %>%  # Remove missing values
+  group_by(`Study Arm`, Outcome, Present) %>%
+  summarise(Total = n(), .groups = "drop")
 
-  # Display only the formatted table
-  baseline_characteristics
-} else {
-  cat("️ No data available for analysis.")
-}
+# ✅ Convert outcome values to categorical (if numeric)
+pregnancy_data$Present <- factor(pregnancy_data$Present, levels = c(0, 1), labels = c("Absent", "Present"))
 
-```
-
-```{r, echo=FALSE, message=FALSE, warning=FALSE, fig.show="hold"}
-# Load required libraries
-library(dplyr)
-library(gtsummary)
-
-# Define malaria outcome variables
-malaria_outcome_vars <- c(
-  "total_malaria_episodes_during_pregnancy",
-  "placental_malaria",
-  "preterm_birth_ontoneo",
-  "stillbirth",
-  "infant_low_birth_weight_2500_g"
-)
-
-# Ensure that `study_arm` is in the dataset
-if (!"study_arm" %in% colnames(PROMO_Data)) {
-  stop("⚠️ Error: `study_arm` column not found in PROMO_Data. Please check column names.")
-}
-
-# Remove `stillbirth` and any other variable that has only one level
-valid_outcomes <- malaria_outcome_vars[sapply(PROMO_Data[malaria_outcome_vars], function(x) length(unique(na.omit(x))) > 1)]
-
-# Ensure `study_arm` is included in the selected variables
-valid_outcomes <- c(valid_outcomes, "study_arm")
-
-# Define label mappings dynamically based on valid variables
-all_labels <- list(
-  total_malaria_episodes_during_pregnancy ~ "Total Malaria Episodes (Mean ± SD)",
-  placental_malaria ~ "Placental Malaria (Histology)",
-  preterm_birth_ontoneo ~ "Preterm Birth",
-  stillbirth ~ "Stillbirth",
-  infant_low_birth_weight_2500_g ~ "Low Birth Weight (<2500g)"
-)
-
-dynamic_labels <- all_labels[names(all_labels) %in% valid_outcomes]
-
-# Define a custom function to format p-values
-format_pvalue <- function(pvals) {
-  sapply(pvals, function(x) {
-    if (is.na(x)) {
-      return("NA")  # Handle missing values properly
-    } else if (is.numeric(x)) {
-      if (x < 0.001) {
-        return("< 0.001")  # Report very small values as "< 0.001"
-      } else if (x < 0.01) {
-        return(format(round(x, 3), nsmall = 3))  # Report between 0.001 and 0.01 to three decimal places
-      } else {
-        return(format(round(x, 2), nsmall = 2))  # Report >= 0.01 to two decimal places
-      }
-    } else {
-      return("NA")  # Return "NA" for non-numeric values
-    }
-  })
-}
-
-# Generate table with formatted p-values
-if (nrow(PROMO_Data) > 0 && length(valid_outcomes) > 1) {
-
-  malaria_outcomes_table <- PROMO_Data %>%
-    select(all_of(valid_outcomes)) %>%
-    tbl_summary(
-      by = "study_arm",
-      statistic = list(
-        total_malaria_episodes_during_pregnancy ~ "{mean} ± {sd}",
-        all_categorical() ~ "{n} ({p}%)"
-      ),
-      label = dynamic_labels
-    ) %>%
-    add_p(
-      test = list(
-        all_categorical() ~ "fisher.test",
-        all_continuous() ~ "t.test"
-      ),
-      pvalue_fun = format_pvalue  # Apply the fixed p-value formatting function
-    )
-
-  # Print the updated table
-  print(malaria_outcomes_table)
-
-} else {
-  message("⚠️ No data available for analysis or insufficient variability in variables.")
-}
+# ✅ Plot Pregnancy Outcomes by Study Arm
+ggplot(pregnancy_data, aes(x = Outcome, y = Total, fill = `Study Arm`)) +
+  geom_bar(stat = "identity", position = "dodge") +
+  labs(title = "Pregnancy Outcomes",
+       x = "Pregnancy Outcome",
+       y = "Total Count",
+       fill = "Study Arm") +
+  theme_minimal() +
+  scale_fill_manual(values = c("lightblue", "pink")) +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
 
